### 2024-01-23 outputs/paper/paper.qmd
change 'XX' and 'None' to NA for easier data processing

---
title: "An Intersectional Analysis of Police Arrests "
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

---
title: "Do More Subdivisions in Wards Increase Voter Turnout? An Analysis of the 2022 Toronto Municipal Election"
author: "Inessa De Angelis"
thanks: "Code and data from this analysis are available at: https://github.com/InessaDeAngelis/Toronto_Elections" 
date: "today"
date-format: "long" 
abstract: "The 2022 Toronto Municipal Election witnessed the lowest voter turnout in the city’s history since amalgamation in 1997, with voter turnout on a ward-by-ward basis ranging from 22% to 38%. Using data from OpenDataToronto, this paper disproves the hypothesis that having more subdivisions within a ward bolsters voter turnout. This paper reveals that more subdivisions within a ward does not necessarily correlate to a higher voter turnout."
format: pdf
toc: true
number-sections: true
bibliography: ref.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-01-23 scripts/02-data_cleaning.R

@@ -48,9 +48,9 @@ cleaned_arrests_strip_data <-
                                  reason_weapons = c("None"),
                                  reason_has_evidence = c("None")))
 
-# Replace all "XX" and "None" from the columns arrest_loc_div and items_found with NA and 0 respectively to make the data processing easier later on
+# Replace all "XX" and "None" from the columns arrest_loc_div and items_found with NA and NA respectively to make the data processing more efficient later on
 # Code referenced from https://www.youtube.com/watch?v=g1eppE60J5g
-cleaned_arrests_strip_data$items_found[cleaned_arrests_strip_data$items_found == "None"] <- 0
+cleaned_arrests_strip_data$items_found[cleaned_arrests_strip_data$items_found == "None"] <- NA
 cleaned_arrests_strip_data$arrest_loc_div[cleaned_arrests_strip_data$arrest_loc_div == "XX"] <- NA
 
 

### 2024-01-23 scripts/00-simulate_data.R
add set.seed(56)

@@ -19,6 +19,8 @@ library(janitor)
 #### Simulate the victim's gender, racial identity and the strip-search nature of each police arrest event ####
 # Code referenced from from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#simulate
 
+set.seed(56) #ensure simulated data's reproducibility
+
 simulated_data <-
   tibble(
     #use 1 through 1000 to represent each event
@@ -47,27 +49,39 @@ simulated_data
 
 #### Summarize the victim's gender identity of each police arrest event ####
 
+set.seed(56)
+
 gender_per_event = 
   simulated_data |>
   group_by(sex, strip_search) |>
   count() |>
   rename(number_of_arrests = n)
+
+gender_per_event
   
 #### Summarize the victim's racial identity of each police arrest event ####
+
+set.seed(56)
 
 race_per_event = 
   simulated_data |>
   group_by(race, strip_search) |>
   count() |>
   rename(number_of_arrests = n)
+
+race_per_event
   
 #### Summarize whether each police arrest event contains strip search ####
+
+set.seed(56)
 
 strip_search_per_event = 
   simulated_data |>
   group_by(strip_search) |>
   count() |>
   rename(number_of_arrests = n)
+
+strip_search_per_event
 
 
 #### Create histograms and graphs ####
@@ -108,7 +122,6 @@ gender_per_event |>
 
 
 
-
 #### Data Validation ####
 
 # Code referenced from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
@@ -130,3 +143,8 @@ simulated_data$race |> unique() |> length() == 9
 
 # Check that there are exactly 3 gender identities identified #
 simulated_data$sex |> unique() |> length() == 3
+
+# Check that there are exactly 2 options for strip searches #
+simulated_data$strip_search |> unique() |> length() == 2
+
+

### 2024-01-23 scripts/02-data_cleaning.R
update clarifying comments on relocate()

@@ -42,7 +42,7 @@ cleaned_arrests_strip_data <-
          race = perceived_race,
          age_group = age_group_at_arrest
          ) |> 
-  relocate(event_id) |>
+  relocate(event_id) |> # Put the event_id column to be the first one in the dataset
   naniar::replace_with_na(replace = list(reason_injury = c("None"), # Replace the string 'None' with NA to make it easier to manipulate data later on
                                  reason_escape = c("None"),
                                  reason_weapons = c("None"),

### 2024-01-23 scripts/02-data_cleaning.R
change a variable name

@@ -40,7 +40,8 @@ cleaned_arrests_strip_data <-
          reason_weapons = search_reason_possess_weapons,
          reason_has_evidence = search_reason_possess_evidence,
          race = perceived_race,
-         age_group = age_group_at_arrest
+         age_group = age_group_at_arrest,
+         gender = sex
          ) |> 
   relocate(event_id) |> # Put the event_id column to be the first one in the dataset
   naniar::replace_with_na(replace = list(reason_injury = c("None"), # Replace the string 'None' with NA to make it easier to manipulate data later on

### 2024-01-23 outputs/paper/paper.qmd
refine abstract and introduction

@@
 ---
-title: "An Intersectional Analysis of Police Arrests "
-subtitle: "My subtitle if needed"
-author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+title: "An Alarming Rate of Gendered and Racialized Police Strip Searches in the Toronto Municipal, 2020-2021"
+author: "Quang Mai"
+thanks: "Code and data are available at: https://github.com/ponolite/Police_Arrests_and_Strip_Searches/tree/main"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Police enforcement, particularly arrests and strip searches, has been a topic of great contention during the COVID-19 pandemic.
+The question commonly raised was how certain demographics are more likely to undergo arrests and strip searches than others, and why has this skewed power dynamics been maintained for so long.
+Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021.
+This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities." WHAT WAS FOUND!!
 format: pdf
@@
 bibliography: references.bib
----
-
----
-title: "Do More Subdivisions in Wards Increase Voter Turnout? An Analysis of the 2022 Toronto Municipal Election"
-author: "Inessa De Angelis"
-thanks: "Code and data from this analysis are available at: https://github.com/InessaDeAngelis/Toronto_Elections" 
-date: "today"
-date-format: "long" 
-abstract: "The 2022 Toronto Municipal Election witnessed the lowest voter turnout in the city’s history since amalgamation in 1997, with voter turnout on a ward-by-ward basis ranging from 22% to 38%.
-Using data from OpenDataToronto, this paper disproves the hypothesis that having more subdivisions within a ward bolsters voter turnout.
-This paper reveals that more subdivisions within a ward does not necessarily correlate to a higher voter turnout."
-format: pdf
-toc: true
-number-sections: true
-bibliography: ref.bib
 ---
@@
 library(tidyverse)
-library(palmerpenguins)
+library(knitr)
+library(opendatatoronto)
 ```
@@
 
-You can and should cross-reference sections and sub-sections.
+In recent times, the Ford government has been at the forefront of public push-back, due to their revisions of strip search laws in Toronto under the Ministry of Correctional Services Act, R.S.O.
+1990, c.
+M.22.
+Addressed by the Canadian Civil Liberties Association, the revisions "fall short of what is needed to ensure that prisoners are treated with basic respect" (@ontarioStripLaw).
+By definition, strip search is a police enforcement, which entails removing some or all articles of clothing to visually inspect a person's body and see if they harbor any items like weapons or drugs (@torontoPolice).
+Long considered one of the most dehumanizing policing acts, strip search often requires 'suspects' to bend over to have their private parts examined.
+Surprisingly, while contended as an alarming issue, little has been done to contribute to the legal and political conversations around strip searches as a form of police power abuse in Toronto.
+This is especially relevant considering the city's racialized and gendered demographics, and how strip searches most likely affect these social groups (@stripSearchOpEd).
+It's evermore relevant looking back at strip searches under the context of the global pandemic COVID-19 during 2020 and 2021, where much of Toronto's diverse population is left vulnerable when interacting with police forces (@covidStripSearch).
 
-The remainder of this paper is structured as follows. @sec-data....
+This paper thus leverages gender and racial identities of 'suspects' arrested by Toronto police from 2020 to 2021, and assesses if there is a correlation between racialized and gendered Toronto residents and strip searching.
+The converging social identities of gender and race may prove to be a empowering source to the 'suspects', but a factor of oppression to the police forces.
+Therefore, this paper aims to contribute to the conversation around strip searching of the vulnerable population, and aims to be a form of data activism to counter to injustices related to police power abuse under the subtext of COVID-19.
+To accordingly analyze, I first map out the distribution of Toronto police arrests by race and gender from 2020 to 2021 using bar graphs and time plots.
+Then with each graph, and for each variable of race and gender, the distribution bars would be sectioned into including strip searches or not to visualize the statistics for each racial and gender identities identified.
+With that,
+
+analyze the distribution of gender and racial identities affected by strip searches [CONCLUSION!] To accordingly come to this conclusion,
+
+Structurally, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
+First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data cleaning and analysis took place (@torontoPolice).
+Then, all the trends and findings I discovered are shown in the Results section, while the Discussion section further assesses these findings.
+Ultimately, the Conclusion section summarizes the paper's main insights.
+In additon, it's worth noting that the data was sourced from Open Data Toronto and Toronto Police Services themselves.
+Thus, this dataset may already contain much racial and gender bias from the very source that this paper is questioning: the Toronto Police Services and its enforcement of strip searching (@stripSearchOpEd).
+Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021, yet it's still worth looking at due the alarming precedents of police power abuse.
 
@@
 Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
 

### 2024-01-23 outputs/paper/paper.qmd
refine abstract

@@
 Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021.
-This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities." WHAT WAS FOUND!!
+This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities."
 format: pdf
@@
 
-# Introduction
+## Introduction
 
@@
 
-Structurally, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
+Howevever, it's worth noting that the data was sourced from Open Data Toronto and Toronto Police Services themselves.
+Thus, this dataset may already contain much racial and gender bias from the very source that this paper is questioning: the Toronto Police Services and its enforcement of strip searching (@stripSearchOpEd).
+Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021, yet it's still worth looking at due the alarming precedents of police power abuse.
+Nonetheless, structurally, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
 First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data cleaning and analysis took place (@torontoPolice).
@@
 Ultimately, the Conclusion section summarizes the paper's main insights.
-In additon, it's worth noting that the data was sourced from Open Data Toronto and Toronto Police Services themselves.
-Thus, this dataset may already contain much racial and gender bias from the very source that this paper is questioning: the Toronto Police Services and its enforcement of strip searching (@stripSearchOpEd).
-Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021, yet it's still worth looking at due the alarming precedents of police power abuse.
 
-# Data {#sec-data}
+## Data
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+The data package used for analysis was sourced from the Open Data Toronto Portal under the library `opendatatoronto` [@rOpenDataToronto].
+Only one dataset was retrieved to examine the racial and gender identities of Toronto police arrests from 2020 to 2021: `Police Race and Identity Based Data - Arrests and Strip Searches` [@torontoPolice].
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra], `naniar` [@rNaniar] and `knitr` [@rKnitr].
+Additional details on data extraction and cleaning processes will be specified in subsequent sections below.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
 #| echo: false
+#| message: false
+#| label: tbl-cleaned_clinic_data
+#| tbl-cap: Sample of cleaned Covid-19 clinic data
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+#referenced kable stying from https://github.com/EthanSansom/torontocitytrees/blob/main/outputs/paper/paper.Rmd
+head(cleaned_clinic_data, 5) |>
+  kable(
+    col.names = c("ID", "Name", "Type", "Ward Name"),
+    booktabs = TRUE
+  )
 ```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
@@
 
-# Results
+## Results
 
@@
 
-# Discussion
+## Discussion
 
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+## Conclusion
 
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
-\newpage
-
-# References
+## References

### 2024-01-23 scripts/02-data_cleaning.R
further cleaning data

@@ -23,6 +23,7 @@ library(naniar)
 #### Clean data ####
 # Code of function replace_with_na referenced from https://tidyr.tidyverse.org/reference/replace_na.html
 
+# Clean from the raw dataset
 raw_police_arrests_strip_data <- read_csv("inputs/data/raw_police_arrests_strip_data.csv")
 
 cleaned_arrests_strip_data <-
@@ -54,6 +55,20 @@ cleaned_arrests_strip_data <-
 cleaned_arrests_strip_data$items_found[cleaned_arrests_strip_data$items_found == "None"] <- NA
 cleaned_arrests_strip_data$arrest_loc_div[cleaned_arrests_strip_data$arrest_loc_div == "XX"] <- NA
 
+# Clean from the cleaned dataset, to keep the a dataset of variables like event_id, race, gender, age_group, strip search and items_found
+
+cleaned_race_gender_data = 
+  cleaned_arrests_strip_data |>
+  select(event_id, race, gender, age_group, strip_search, items_found)
+
+# Clean from the cleaned dataset, to keep a dataset of variables like reasons for strip searching
+
+cleaned_search_reasons_data = 
+  cleaned_arrests_strip_data |>
+  select(event_id, reason_injury,	reason_escape, reason_weapons, reason_has_evidence, strip_search, items_found) |>
+  filter(!is.na(reason_injury) | !is.na(reason_escape) | !is.na(reason_weapons) | !is.na(reason_has_evidence))
 
 #### Save data ####
 write_csv(cleaned_arrests_strip_data,"inputs/data/cleaned_arrests_strip_data.csv")
+write_csv(cleaned_race_gender_data,"inputs/data/cleaned_race_gender_data.csv")
+write_csv(cleaned_search_reasons_data,"inputs/data/cleaned_search_reasons_data.csv")

### 2024-01-23 scripts/00-simulate_data.R
match simulation with raw data

@@ -23,24 +23,24 @@ set.seed(56) #ensure simulated data's reproducibility
 
 simulated_data <-
   tibble(
-    #use 1 through 1000 to represent each event
-    "event" = 1:1000,
-    #randomly pick a race identified from the dataset, with replacement, 1000 times
+    #use 1 through 65276 to represent each event
+    "event" = 1:65276,
+    #randomly pick a race identified from the dataset, with replacement, 65276 times
     "race" = sample(
       x = c("White", "Unknown or Legacy", "Black", "South Asian", "Indigenous", "Middle-Eastern", "Latino", "East/Southeast Asian", "None"),
-      size = 1000,
+      size = 65276,
       replace = TRUE
     ),
-    #randomly pick one gender identity identified from the dataset, with replacement, 1000 times
+    #randomly pick one gender identity identified from the dataset, with replacement, 65276 times
     "sex" = sample(
       x = c("Male", "Female", "Unidentified"),
-      size = 1000,
+      size = 65276,
       replace = TRUE
     ),
-    #randomly pick whether the event includes strip search, with replacement, 1000 times
+    #randomly pick whether the event includes strip search, with replacement, 65276 times
     "strip_search" = sample(
       x = c("Yes", "No"),
-      size = 1000,
+      size = 65276,
       replace = TRUE
     )
 )
@@ -126,11 +126,11 @@ gender_per_event |>
 
 # Code referenced from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
 
-# Check that there are between 1 and 1000 events of police arrests and strip searches #
+# Check that there are between 1 and 65276 events of police arrests and strip searches #
 simulated_data$event |> min() == 1
-simulated_data$event |> max() == 1000
+simulated_data$event |> max() == 65276
 
-# Check that there are no more than 1000 events #
+# Check that there are no more than 65276 events #
 simulated_data |>
   group_by(event) |>
   count() |>

### 2024-01-23 scripts/00-simulate_data.R
polish simulation and data cleaning

@@ -47,6 +47,21 @@ simulated_data <-
 
 simulated_data
 
+#add a new items_found column to those events with strip search, with replacement, 32717 times (32717 is based on the set.seed() number)
+
+set.seed(56)
+
+simulated_data_items <-
+  simulated_data |>
+  filter(strip_search == "Yes") |>
+  mutate(
+    "items_found" = sample(
+      x = c("Yes", "No"),
+      size = 32717,
+      replace = TRUE
+    )
+  )
+
 #### Summarize the victim's gender identity of each police arrest event ####
 
 set.seed(56)

### 2024-01-23 scripts/02-data_cleaning.R

@@ -59,7 +59,7 @@ cleaned_arrests_strip_data$arrest_loc_div[cleaned_arrests_strip_data$arrest_loc_
 
 cleaned_race_gender_data = 
   cleaned_arrests_strip_data |>
-  select(event_id, race, gender, age_group, strip_search, items_found)
+  select(event_id, race, gender, strip_search, items_found)
 
 # Clean from the cleaned dataset, to keep a dataset of variables like reasons for strip searching
 

### 2024-01-24 scripts/02-data_cleaning.R
update data cleaning and naming

@@ -39,7 +39,7 @@ cleaned_arrests_strip_data <-
          reason_injury = search_reason_cause_injury,
          reason_escape = search_reason_assist_escape,
          reason_weapons = search_reason_possess_weapons,
-         reason_has_evidence = search_reason_possess_evidence,
+         reason_possess_evidence = search_reason_possess_evidence,
          race = perceived_race,
          age_group = age_group_at_arrest,
          gender = sex
@@ -65,8 +65,13 @@ cleaned_race_gender_data =
 
 cleaned_search_reasons_data = 
   cleaned_arrests_strip_data |>
-  select(event_id, reason_injury,	reason_escape, reason_weapons, reason_has_evidence, strip_search, items_found) |>
-  filter(!is.na(reason_injury) | !is.na(reason_escape) | !is.na(reason_weapons) | !is.na(reason_has_evidence))
+  select(event_id, reason_injury,	reason_escape, reason_weapons, reason_possess_evidence) |>
+  filter(!is.na(reason_injury))
+
+
+  cleaned_arrests_strip_data |>
+  select(event_id, reason_injury,	reason_escape, reason_weapons, reason_possess_evidence, items_found) |>
+  filter(reason_injury == 0 & reason_escape == 0 & reason_weapons == 0 & reason_possess_evidence == 0)
 
 #### Save data ####
 write_csv(cleaned_arrests_strip_data,"inputs/data/cleaned_arrests_strip_data.csv")

### 2024-01-24 outputs/paper/paper.qmd
complete data section

@@
 
-In recent times, the Ford government has been at the forefront of public push-back, due to their revisions of strip search laws in Toronto under the Ministry of Correctional Services Act, R.S.O.
+In recent times, the Ford government has been at the forefront of public push-back in Toronto, due to their revisions of strip search laws under the Ministry of Correctional Services Act, R.S.O.
 1990, c.
@@
 This is especially relevant considering the city's racialized and gendered demographics, and how strip searches most likely affect these social groups (@stripSearchOpEd).
-It's evermore relevant looking back at strip searches under the context of the global pandemic COVID-19 during 2020 and 2021, where much of Toronto's diverse population is left vulnerable when interacting with police forces (@covidStripSearch).
+It's evermore relevant looking back at strip searches under the context of the global pandemic COVID-19 during 2020 and 2021, where much of Toronto's diverse, racialized population is left vulnerable when interacting with police forces (@covidStripSearch).
 
@@
 Therefore, this paper aims to contribute to the conversation around strip searching of the vulnerable population, and aims to be a form of data activism to counter to injustices related to police power abuse under the subtext of COVID-19.
-To accordingly analyze, I first map out the distribution of Toronto police arrests by race and gender from 2020 to 2021 using bar graphs and time plots.
-Then with each graph, and for each variable of race and gender, the distribution bars would be sectioned into including strip searches or not to visualize the statistics for each racial and gender identities identified.
+To accordingly analyze, I first map out the distribution of Toronto police arrests by race and gender from 2020 to 2021 using bar graphs and various other graphs Then with each graph, and for each variable of race and gender, the distribution bars would be sectioned into 'including strip searches' or 'not including' to visualize the statistics of strip searching for each racial and gender identity identified.
 With that,
@@
 The data package used for analysis was sourced from the Open Data Toronto Portal under the library `opendatatoronto` [@rOpenDataToronto].
-Only one dataset was retrieved to examine the racial and gender identities of Toronto police arrests from 2020 to 2021: `Police Race and Identity Based Data - Arrests and Strip Searches` [@torontoPolice].
+Only one dataset was retrieved from the data package to examine the racial and gender identities of Toronto police arrests from 2020 to 2021, which is the package `Police Race and Identity Based Data - Arrests and Strip Searches` [@torontoPolice].
 Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra], `naniar` [@rNaniar] and `knitr` [@rKnitr].
-Additional details on data extraction and cleaning processes will be specified in subsequent sections below.
+
+Last updated as of December 2, 2022, this chosen dataset was published by Toronto Police Services to convey information related to all arrests and strip searches by their police force from 2020 to 2021.
+Originally, the dataset features information on date of arrest, incident id, arrest id, person id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, whether the person was booked at the police station, the person's action at the time of arrest, and the police's reason for strip searching along with details on any items found.
+While the dataset contains a wide range of information, there was much naming inconsistency and many variables outside the scope of this paper.
+Thus, in order to conduct an intersectional—or gender and race—analysis of police strip searches, I have to further clean this dataset and split it into two smaller datasets.
+Additional details on data extraction and cleaning will be specified in subsequent sections below.
+
+### Racial and Gender Identities of Police Strip Search
+
+To streamline the data regarding racial and gender identities of the people involved in police strip searches, the names within the original dataset are simplified and then the dataset is split into a smaller one containing relevant columns (see @tbl-cleaned_race_gender_data).
+This dataset contains demographic and strip search-relevant information for each police arrest event, such as race, gender, whether the event involves a strip search, and if it is a strip search, are there any items found.
+
+Particularly, the column 'Strip Search' signifies whether an arrest involves strip search.
+If the data cell contains 'NA', this means that the event in particular doesn't have a strip search.
+But if it is a '1' in that particular data cell, this means that the event does involve involve a strip search.
+This goes the same  for the 'Items Found' column, where '1s' and '0s' will only appear if the 'Strip Search' value of an event is a '1'.
+This is because the items found within a strip search is only available when a police arrest does involve strip searching.
+Moreover, retainment of the numeric '1' and '0' instead of 'Yes' or 'No' ensures that the data can be quantified later one.
+
+Furthermore, the reason for keeping rows of events that don't include strip search is to later produce summary statistics of events with and without strip searches.
+This further aligns with the task of the paper at hand, which is to critically reflect on the rate of strip searching on people with racialized and gendered identities.
 
@@
 #| message: false
-#| label: tbl-cleaned_clinic_data
-#| tbl-cap: Sample of cleaned Covid-19 clinic data
+#| label: tbl-cleaned_race_gender_data
+#| tbl-cap: Sample of Cleaned Race and Gender Data against Police Arrests with Strip Searches
 
 #referenced kable stying from https://github.com/EthanSansom/torontocitytrees/blob/main/outputs/paper/paper.Rmd
-head(cleaned_clinic_data, 5) |>
+head(cleaned_race_gender_data, 5) |>
   kable(
-    col.names = c("ID", "Name", "Type", "Ward Name"),
+    col.names = c("ID", "Race", "Gender", "Strip Search", "Items Found"),
     booktabs = TRUE
@@
 ```
+
+### Reasons for Police Strip Search
+
+The other dataset that the original dataset is split into conveys all police arrests that involve strip searching and the reasons behind strip searching.
+Under close examination, the naming of individual data cells in different columns from the original, raw dataset is quite disordered, for instance, using 'None' and 'XX' in place of the usual 'NA'.
+Thus the all inconsistent data cells were renamed to streamline the naming conventions and data processing later on.
+After renaming, the big dataset was split into containing relevant columns (see @tbl-cleaned_search_reasons_data).
+
+Except for "ID" column, all other columns in this new dataset are all reasons which substantiate why the Toronto police force conducts  strip search under certain arrests.
+"Reason: Injury" is for people who have caused injuries, "Reason: Assist Escape" is for people who may have helped other suspects escape, "Reason: Weapons" is for people who possess weapons, and "Reason: Possess Evidence" is for people who are suspected of harboring items like drugs or an alibi.
+For each "Reason" colum, if the data cell contains '0', this means that strip search doens't happen because of that particular reason.
+Vice versa, if it contains a '1', this means strip search does happen because of that reason.
+For each police arrest event in this dataset, there can be none to multiple reasons combined which lead to strip searching.
+
+Then, for this particular dataset that only retains reasons for strip searching, all instances of police arrests without strip searching are dropped using the 'filter()' function, along an '!is.na()' function of any "reason" column nested within.
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-cleaned_search_reasons_data
+#| tbl-cap: Sample of Cleaned Reasons for Police Strip Searches Data
+
+#referenced kable stying from https://github.com/EthanSansom/torontocitytrees/blob/main/outputs/paper/paper.Rmd
+head(cleaned_search_reasons_data, 5) |>
+  kable(
+    col.names = c("ID", "Reason - Injury", "Reason - Escape", "Reason - Weapons", "Reason - Has Evidence"),
+    booktabs = TRUE
+  )
+```
+
+## Results
 
@@
 Talk way more about it.
-
-## Results
 

### 2024-01-24 outputs/paper/paper.qmd
update new data and further clean data

--- 2024-01-23 outputs/paper/paper.qmd
@@
 ---
-title: "An Alarming Rate of Gendered and Racialized Police Strip Searches in the Toronto Municipal, 2020-2021"
+title: "Slight Positive Correlation Between Gendered and Racialized Police Strip Searches in the Toronto Municipal, 2020-2021"
+title-block-banner: true
+abstract: "Police strip searche has been a topic of great contention during the COVID-19 pandemic.
+The question commonly raised was how certain demographics are more likely to undergo arrests and strip searches than others, and why has this skewed power dynamics been maintained for so long.
+Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021.
+This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities.
+More on RESULTS"
+thanks: "Code and data are available at: https://github.com/ponolite/Police_Arrests_and_Strip_Searches/tree/main"
 author: "Quang Mai"
-thanks: "Code and data are available at: https://github.com/ponolite/Police_Arrests_and_Strip_Searches/tree/main"
 date: today
 date-format: long
-abstract: "Police enforcement, particularly arrests and strip searches, has been a topic of great contention during the COVID-19 pandemic.
-The question commonly raised was how certain demographics are more likely to undergo arrests and strip searches than others, and why has this skewed power dynamics been maintained for so long.
-Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021.
-This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities."
 format: pdf
+editor: visual
+toc: true
 number-sections: true
@@
 library(opendatatoronto)
+library(dplyr)
+library(readr)
+library(kableExtra)
+library(here)
+
+#### Read in cleaned data ####
+
+cleaned_arrests_strip_data = read_csv(
+  file = here("inputs/data/cleaned_arrests_strip_data.csv"),
+  show_col_types = FALSE
+)
+
+cleaned_race_gender_data = read_csv(
+  file = here("inputs/data/cleaned_race_gender_data.csv"),
+  show_col_types = FALSE
+)
+
+cleaned_search_reasons_data = read_csv(
+  file = here("inputs/data/cleaned_search_reasons_data.csv"),
+  show_col_types = FALSE
+)
 ```
@@
 
-In recent times, the Ford government has been at the forefront of public push-back, due to their revisions of strip search laws in Toronto under the Ministry of Correctional Services Act, R.S.O.
+Recently, the Ford government has been at the forefront of public push-back in Toronto, due to their revisions of strip search laws under the Ministry of Correctional Services Act, R.S.O.
 1990, c.
 M.22.
-Addressed by the Canadian Civil Liberties Association, the revisions "fall short of what is needed to ensure that prisoners are treated with basic respect" (@ontarioStripLaw).
-By definition, strip search is a police enforcement, which entails removing some or all articles of clothing to visually inspect a person's body and see if they harbor any items like weapons or drugs (@torontoPolice).
-Long considered one of the most dehumanizing policing acts, strip search often requires 'suspects' to bend over to have their private parts examined.
-Surprisingly, while contended as an alarming issue, little has been done to contribute to the legal and political conversations around strip searches as a form of police power abuse in Toronto.
-This is especially relevant considering the city's racialized and gendered demographics, and how strip searches most likely affect these social groups (@stripSearchOpEd).
-It's evermore relevant looking back at strip searches under the context of the global pandemic COVID-19 during 2020 and 2021, where much of Toronto's diverse population is left vulnerable when interacting with police forces (@covidStripSearch).
+Addressed by the Canadian Civil Liberties Association, the revisions "fall short...to ensure that prisoners are treated with basic respect" [@ontarioStripLaw].
+By definition, strip search is a police enforcement, which entails removing some or all articles of clothing to visually inspect a person's private body parts and see if they harbor any items like weapons or drugs [@torontoPolice].
+Surprisingly, while contended as a dehumanizing policing act, little has been done to assess strip search as a form of police power abuse in Toronto.
+This is especially relevant considering the city's racialized and gendered demographics, how strip searches most likely affect these social groups under the subtext of COVID-19 from 2020 to 2021 [@stripSearchOpEd].
 
-This paper thus leverages gender and racial identities of 'suspects' arrested by Toronto police from 2020 to 2021, and assesses if there is a correlation between racialized and gendered Toronto residents and strip searching.
-The converging social identities of gender and race may prove to be a empowering source to the 'suspects', but a factor of oppression to the police forces.
-Therefore, this paper aims to contribute to the conversation around strip searching of the vulnerable population, and aims to be a form of data activism to counter to injustices related to police power abuse under the subtext of COVID-19.
-To accordingly analyze, I first map out the distribution of Toronto police arrests by race and gender from 2020 to 2021 using bar graphs and time plots.
-Then with each graph, and for each variable of race and gender, the distribution bars would be sectioned into including strip searches or not to visualize the statistics for each racial and gender identities identified.
-With that,
+This paper thus inspects gender and racial identities of all Toronto police arrests from 2020 to 2021, and assesses if there is a correlation between marginalized identities and police strip searching.
+Foremost, this paper aims to be a form of data activism to counter the injustices related to police power abuse around strip searching of the vulnerable Toronto population.
+To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using bar graphs and various other graphs.
+Particularly, for each variable of race and gender, the distribution bars would be sectioned into 'including strip searches' or 'not including' to visualize the statistics of strip searching.
+With that, It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd].
+Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021.
+Yet, it's still worth looking at due the alarming precedents of police power abuse [@covidStripSearch].
 
-analyze the distribution of gender and racial identities affected by strip searches [CONCLUSION!] To accordingly come to this conclusion,
-
-Howevever, it's worth noting that the data was sourced from Open Data Toronto and Toronto Police Services themselves.
-Thus, this dataset may already contain much racial and gender bias from the very source that this paper is questioning: the Toronto Police Services and its enforcement of strip searching (@stripSearchOpEd).
-Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021, yet it's still worth looking at due the alarming precedents of police power abuse.
-Nonetheless, structurally, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
-First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data cleaning and analysis took place (@torontoPolice).
-Then, all the trends and findings I discovered are shown in the Results section, while the Discussion section further assesses these findings.
+Structurally then, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
+First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data cleaning and extraction took place [@torontoPolice].
+Then, all the trends and findings discovered are shown in the Results section, while the Discussion section further assesses these findings.
 Ultimately, the Conclusion section summarizes the paper's main insights.
@@
 The data package used for analysis was sourced from the Open Data Toronto Portal under the library `opendatatoronto` [@rOpenDataToronto].
-Only one dataset was retrieved to examine the racial and gender identities of Toronto police arrests from 2020 to 2021: `Police Race and Identity Based Data - Arrests and Strip Searches` [@torontoPolice].
+Only one dataset was retrieved from the data package to examine the racial and gender identities of Toronto police arrests from 2020 to 2021, which is the package `Police Race and Identity Based Data - Arrests and Strip Searches` [@torontoPolice].
 Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra], `naniar` [@rNaniar] and `knitr` [@rKnitr].
-Additional details on data extraction and cleaning processes will be specified in subsequent sections below.
+
+### Overview of Data Extraction and Cleaning
+
+This chosen dataset was published by Toronto Police Services to convey information related to all police arrests and strip searches from 2020 to 2021 (last updated on December 2, 2022).
+Originally, the dataset features information on date of arrest, event id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, the person's action at the time of arrest, and the police's reason for strip searching along with details on any items found.
+Informative as it is, the dataset contains much naming inconsistency and many variables outside the scope of this paper.
+Thus, in order to conduct an intersectional---or gender and race---analysis of police strip searches, I have to further clean this dataset and split it into two smaller datasets.
+Additional details on data extraction and cleaning will be specified in subsequent sections.
+
+### Racial and Gender Identities of Police Strip Search
+
+To streamline the data regarding racial and gender identities of the people involved in police strip searches, the names within the original dataset are simplified and then the dataset is split into a smaller one containing relevant columns ([see @tbl-cleaned_race_gender_data]).
+This dataset contains demographic and strip search-relevant information for each police arrest event, such as race, gender, whether the event involves a strip search, and if it is a strip search, are there any items found.
+
+Specifically, the column 'Strip Search' signifies whether an arrest involves strip search.
+As for the 'Items Found' column, either a `1` or a `0` will appear in a data cell if the 'Strip Search' value of an event is a `1`.
+This is because the items found within a strip search is only available when a police arrest does involve strip searching.
+
+Furthermore, the reason for keeping rows of events that don't include strip search is to later produce summary statistics of events with and without strip searches.
+This further aligns with the task of the paper at hand, which is to critically reflect on the rate of strip searching on people with racialized and gendered identities.
 
@@
 #| message: false
-#| label: tbl-cleaned_clinic_data
-#| tbl-cap: Sample of cleaned Covid-19 clinic data
+#| label: tbl-cleaned_race_gender_data
+#| tbl-cap: Sample of Race and Gender Data against Police Arrests with Strip Searches
 
 #referenced kable stying from https://github.com/EthanSansom/torontocitytrees/blob/main/outputs/paper/paper.Rmd
-head(cleaned_clinic_data, 5) |>
+head(cleaned_race_gender_data, 12) |>
   kable(
-    col.names = c("ID", "Name", "Type", "Ward Name"),
+    col.names = c("ID", "Race", "Gender", "Strip Search", "Items Found"),
+    booktabs = TRUE
+  )
+```
+
+### Reasons for Police Strip Search
+
+The other dataset conveys all police arrests that involve strip searching and the reasons behind strip searching.
+Under close examination, the naming of individual data cells in different columns from the original, raw dataset is quite disordered, for instance, using `None` and `XX` in place of the usual `NA`.
+Thus, all inconsistent data cells were renamed to streamline the naming conventions and data processing later on.
+After renaming, the big dataset was split into containing relevant columns ([see @tbl-cleaned_search_reasons_data]).
+Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for rows with `NA` as values all the `reason` column.
+
+Except for "ID" column, all other columns in this new dataset are reasons which account for why the Toronto police force conducts strip search under certain arrests.
+For instance, "Reason - Assist Escape" is for people who may have helped other suspects escape and "Reason - Possess Evidence" is for people who are suspected of harboring items like drugs or an alibi.
+For each "Reason" column, if the data cell contains `0`, this means that strip search doesn't happen because of that particular reason.
+Vice versa, if it contains `1`, this means strip search happens because of that reason.
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-cleaned_search_reasons_data
+#| tbl-cap: Sample of Reasons for Police Strip Searches Data
+
+head(cleaned_search_reasons_data, 12) |>
+  kable(
+    col.names = c("ID", "Reason - Injury", "Reason - Escape", "Reason - Weapons", "Reason - Has Evidence"),
+    booktabs = TRUE
+  )
+```
+
+## Results
+
+### Racialized and Gendered Police Strip Searches Statistics
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-race-strip-search
+#| tbl-cap: Summary Statistics of Racialized Police Strip Searches
+
+cleaned_race_gender_data |>
+  group_by(race, strip_search) |>
+  count() |>
+  rename(number_of_arrests = n) |>
+  group_by(race) |>
+  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
+  kable(
+    col.names = c("Race", "Strip Search", "Number of Arrests", "Percentage"),
+    booktabs = TRUE)
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| label: fig-race_strip_search
+#| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
+
+  cleaned_race_gender_data |>
+  group_by(race, strip_search) |>
+  count() |>
+  rename(number_of_arrests = n) |>
+  ggplot(aes(x = race, y = number_of_arrests, fill = strip_search)) + 
+  geom_bar(stat="identity") + 
+  theme_minimal() +
+  theme(legend.key.size = unit(0.5, 'cm'), axis.text.x = element_text(angle = 70, vjust = 1, hjust= 1)) +
+  labs(x = "Race",
+       y = "Number of Arrests",
+       fill = "Strip Search"
+  ) +
+  scale_fill_viridis_d(option="cividis")
+
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-gender-strip-search
+#| tbl-cap: Summary Statistics of Gendered Police Strip Searches
+
+cleaned_race_gender_data |>
+  group_by(gender, strip_search) |>
+  count() |>
+  rename(number_of_arrests = n) |>
+  group_by(gender) |>
+  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
+  kable(
+    col.names = c("Gender", "Strip Search", "Number of Arrests", "Percentage"),
+    booktabs = TRUE)
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| label: fig-gender_strip_search
+#| fig-cap: Gender Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
+
+  cleaned_race_gender_data |>
+  group_by(gender, strip_search) |>
+  count() |>
+  rename(number_of_arrests = n) |>
+  ggplot(aes(x = gender, y = number_of_arrests, fill = strip_search)) + 
+  geom_bar(stat="identity") + 
+  theme_minimal() +
+  theme(legend.key.size = unit(0.5, 'cm')) +
+  labs(x = "Gender",
+       y = "Number of Arrests",
+       fill = "Strip Search"
+  ) +
+  scale_fill_viridis_d(option="cividis")
+
+```
+### Police Strip Searches and Items Found Statistics
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-items-strip-search
+#| tbl-cap: Summary Statistics of Items Found During Police Strip Searches
+
+cleaned_race_gender_data |>
+  group_by(strip_search, items_found) |>
+  count() |>
+  rename(number_of_arrests = n) |>
+  filter(strip_search == "Yes") |>
+  select(items_found, number_of_arrests) |>
+  kable(
+    col.names = c("Strip Search", "Items Found", "Number of Arrests"),
     booktabs = TRUE
@@
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
 #| echo: false
-#| warning: false
+#| include: false
 #| message: false
-
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-## Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+  cleaned_arrests_strip_data |>
+  select(event_id, reason_injury,	reason_escape, reason_weapons, reason_possess_evidence, items_found) |>
+  filter(reason_injury == 0 & reason_escape == 0 & reason_weapons == 0 & reason_possess_evidence == 0)
 ```

### 2024-01-24 scripts/02-data_cleaning.R

@@ -29,14 +29,9 @@ raw_police_arrests_strip_data <- read_csv("inputs/data/raw_police_arrests_strip_
 cleaned_arrests_strip_data <-
   raw_police_arrests_strip_data |>
   janitor::clean_names() |> #clean and simplify variable names 
-  select(-c(x_id, arrest_id, person_id, object_id, booked)) |> # Get rid of unnecessary variables
-  rename(action_concealed = actions_at_arrest_concealed_i, # Shorten variable names to make them readable
-         action_combative = actions_at_arrest_combative,
-         action_resisted = actions_at_arrest_resisted_d,
-         action_mental_inst = actions_at_arrest_mental_inst,
-         action_assaulted = actions_at_arrest_assaulted_o,
-         action_cooperative = actions_at_arrest_cooperative,
-         reason_injury = search_reason_cause_injury,
+  select(-c(x_id, arrest_id, person_id, object_id, booked, actions_at_arrest_concealed_i, actions_at_arrest_combative, actions_at_arrest_resisted_d, actions_at_arrest_mental_inst, actions_at_arrest_assaulted_o, actions_at_arrest_cooperative)) |> # Get rid of unnecessary variables
+  rename(
+         reason_injury = search_reason_cause_injury, # Shorten variable names to make them readable
          reason_escape = search_reason_assist_escape,
          reason_weapons = search_reason_possess_weapons,
          reason_possess_evidence = search_reason_possess_evidence,
@@ -48,9 +43,9 @@ cleaned_arrests_strip_data <-
   naniar::replace_with_na(replace = list(reason_injury = c("None"), # Replace the string 'None' with NA to make it easier to manipulate data later on
                                  reason_escape = c("None"),
                                  reason_weapons = c("None"),
-                                 reason_has_evidence = c("None")))
+                                 reason_possess_evidence = c("None")))
 
-# Replace all "XX" and "None" from the columns arrest_loc_div and items_found with NA and NA respectively to make the data processing more efficient later on
+# Replace all "XX" and "None" from the columns arrest_loc_div and items_found with "NA" respectively to make the data processing more efficient later on
 # Code referenced from https://www.youtube.com/watch?v=g1eppE60J5g
 cleaned_arrests_strip_data$items_found[cleaned_arrests_strip_data$items_found == "None"] <- NA
[…]

### 2024-01-24 outputs/paper/paper.qmd
update and refine data cleaning and results

@@
 title-block-banner: true
-abstract: "Police strip searche has been a topic of great contention during the COVID-19 pandemic.
+abstract: "Police strip search has been a topic of great contention during the COVID-19 pandemic.
 The question commonly raised was how certain demographics are more likely to undergo arrests and strip searches than others, and why has this skewed power dynamics been maintained for so long.
@@
 library(here)
+library(ggbeeswarm)
+library(ggrepel)
+library(ggplot2)
 
@@
 To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using bar graphs and various other graphs.
-Particularly, for each variable of race and gender, the distribution bars would be sectioned into 'including strip searches' or 'not including' to visualize the statistics of strip searching.
+Particularly, for each category of race and gender, the distribution bars would be sectioned into 'including strip searches' or 'not including' to visualize the statistics of strip searching.
 With that, It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd].
 Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021.
-Yet, it's still worth looking at due the alarming precedents of police power abuse [@covidStripSearch].
+Yet, it's still worth looking at due the precedents of police power abuse [@covidStripSearch].
 
@@
 Originally, the dataset features information on date of arrest, event id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, the person's action at the time of arrest, and the police's reason for strip searching along with details on any items found.
-Informative as it is, the dataset contains much naming inconsistency and many variables outside the scope of this paper.
+Informative as it is, the dataset contains much naming inconsistency and many columns outside the scope of this paper.
 Thus, in order to conduct an intersectional---or gender and race---analysis of police strip searches, I have to further clean this dataset and split it into two smaller datasets.
@@
 #referenced kable stying from https://github.com/EthanSansom/torontocitytrees/blob/main/outputs/paper/paper.Rmd
-head(cleaned_race_gender_data, 12) |>
+head(cleaned_race_gender_data, 10) |>
   kable(
@@
 After renaming, the big dataset was split into containing relevant columns ([see @tbl-cleaned_search_reasons_data]).
-Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for rows with `NA` as values all the `reason` column.
+Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for rows with `NA` as values for all the `reason` column.
 
@@
 Vice versa, if it contains `1`, this means strip search happens because of that reason.
+In addition, a new column called "Total Reasons" totals all reasons used to account for each police strip search is added with `mutate()`, where all instances of `1` and `0` are changed into numeric with `as.numeric()`.
 
@@
 
-head(cleaned_search_reasons_data, 12) |>
+  head(cleaned_search_reasons_data, 10) |>
+  select(-c(race, gender)) |>
   kable(
-    col.names = c("ID", "Reason - Injury", "Reason - Escape", "Reason - Weapons", "Reason - Has Evidence"),
+    col.names = c("ID", "Reason - Injury", "Reason - Escape", "Reason - Weapons", "Reason - Has Evidence", "Items_Found", "Total Reasons"),
     booktabs = TRUE
-  )
+  ) |>
+  kable_styling(latex_options="scale_down")
 ```
@@
 ## Results
+
+cite toronto population to say that while the number of white poeple seem large, it's not really
 
@@
 
-```{r}
+```{r fig.height=6, fig.width=7}
 #| echo: false
@@
 
+```{r fig.height=6, fig.width=7}
+#| echo: false
+#| message: false
+#| label: fig-race_strip_search_separate
+#| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
+
+cleaned_race_gender_data |>
+  group_by(race, strip_search) |>
+  count() |>
+  rename(number_of_arrests = n) |>
+  filter(strip_search == "Yes") |> 
+  ggplot(mapping = aes(x = strip_search, fill = strip_search), data=cleaned_race_gender_data) +
+  geom_bar()+theme_minimal()+ylab("number of arrests")+facet_wrap(~race)+ scale_fill_viridis_d(option="cividis")
+
+```
+
 ```{r}
@@
     booktabs = TRUE)
+
 ```
@@
 ```
+
 ### Police Strip Searches and Items Found Statistics
@@
 #| message: false
-#| label: tbl-items-strip-search
+#| label: tbl-reasons-strip-search
 #| tbl-cap: Summary Statistics of Items Found During Police Strip Searches
@@
 #| echo: false
-#| include: false
 #| message: false
-  cleaned_arrests_strip_data |>
-  select(event_id, reason_injury,	reason_escape, reason_weapons, reason_possess_evidence, items_found) |>
-  filter(reason_injury == 0 & reason_escape == 0 & reason_weapons == 0 & reason_possess_evidence == 0)
+#| label: tbl-items-strip-search
+#| tbl-cap: Summary Statistics of Reasons for Police Strip Searches
+
+cleaned_search_reasons_data |>
+  group_by(total_reasons, items_found) |>
+  count() |>
+  rename(count = n) |>
+  kable(
+    col.names = c("Number of Reasons", "Items_Found", "Count"),
+    booktabs = TRUE
+  )
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| label: fig-scatter-plot-reasons
+#| fig-cap: Relationship Between Reasons for and Items Found During Per Strip Search in Toronto, 2020-2021
+
+#create a scatterplot to show each observation in the dataset, comparing the number of reasons to the items found in each strip search
+#colour the points based on the police division where the group of shootings occurred
+#code for geom_point() referenced from https://github.com/TDonofrio62/Shooting-Occurrences-in-Toronto/blob/main/Shooting_Occurrences/outputs/paper/Shooting_Occurrences.Rmd
+cleaned_search_reasons_data |>
+  group_by(total_reasons, items_found, race) |>
+  count() |>
+  rename(count = n) |>
+  ggplot(aes(x = total_reasons, y = count, colour = race)) +
+  geom_point() +
+  theme_bw() +
+  theme(legend.key.size = unit(0.5, 'cm'))+
+  labs(x = "Total Reasons for Strip Searching",
+       y = "Count",
+       color = "Race"
+       ) +
+  scale_fill_viridis_d(option="cividis")
+
+cleaned_search_reasons_data |>
+  group_by(total_reasons, items_found, race) |>
+  count() |>
+  rename(count = n) |>
+  ggplot(aes(x = total_reasons, y = count, colour = race)) +
+  geom_point() +
+  theme_bw() +
+  theme(legend.key.size = unit(0.5, 'cm'))+
+  labs(x = "Total Reasons for Strip Searching",
+       y = "Count",
+       color = "Race"
+       ) +
+  scale_fill_viridis_d(option="cividis") 
 ```

### 2024-01-24 scripts/02-data_cleaning.R

@@ -29,14 +29,13 @@ raw_police_arrests_strip_data <- read_csv("inputs/data/raw_police_arrests_strip_
 cleaned_arrests_strip_data <-
   raw_police_arrests_strip_data |>
   janitor::clean_names() |> #clean and simplify variable names 
-  select(-c(x_id, arrest_id, person_id, object_id, booked, actions_at_arrest_concealed_i, actions_at_arrest_combative, actions_at_arrest_resisted_d, actions_at_arrest_mental_inst, actions_at_arrest_assaulted_o, actions_at_arrest_cooperative)) |> # Get rid of unnecessary variables
+  select(-c(x_id, arrest_id, person_id, object_id, youth_at_arrest_under_18_years, arrest_loc_div, booked, actions_at_arrest_concealed_i, actions_at_arrest_combative, actions_at_arrest_resisted_d, actions_at_arrest_mental_inst, actions_at_arrest_assaulted_o, actions_at_arrest_cooperative)) |> # Get rid of unnecessary variables
   rename(
          reason_injury = search_reason_cause_injury, # Shorten variable names to make them readable
          reason_escape = search_reason_assist_escape,
          reason_weapons = search_reason_possess_weapons,
          reason_possess_evidence = search_reason_possess_evidence,
          race = perceived_race,
-         age_group = age_group_at_arrest,
          gender = sex
          ) |> 
   relocate(event_id) |> # Put the event_id column to be the first one in the dataset
@@ -45,10 +44,13 @@ cleaned_arrests_strip_data <-
                                  reason_weapons = c("None"),
                                  reason_possess_evidence = c("None")))
 
-# Replace all "XX" and "None" from the columns arrest_loc_div and items_found with "NA" respectively to make the data processing more efficient later on
+# Replace all "None" from the column items_found with "NA" to make the data processing more efficient later on
 # Code referenced from https://www.youtube.com/watch?v=g1eppE60J5g
 cleaned_arrests_strip_data$items_found[cleaned_arrests_strip_data$items_found == "None"] <- NA
-cleaned_arrests_strip_data$arrest_loc_div[cleaned_arrests_strip_data$arrest_loc_div == "XX"] <- NA
+
+# Replace all "NA" and "1" from the column strip_search with "No" and "Yes" respectively to make the data clearer later on
+cleaned_arrests_strip_data$strip_search[cleaned_arrests_strip_data$strip_search == 1] <- "Yes"
[…]

### 2024-01-25 scripts/00-simulate_data.R
polish simulated data

@@ -23,7 +23,7 @@ set.seed(56) #ensure simulated data's reproducibility
 
 simulated_data <-
   tibble(
-    #use 1 through 65276 to represent each event
+    #use 1 through 65276 to represent each arrest event
     "event" = 1:65276,
     #randomly pick a race identified from the dataset, with replacement, 65276 times
     "race" = sample(
@@ -32,7 +32,7 @@ simulated_data <-
       replace = TRUE
     ),
     #randomly pick one gender identity identified from the dataset, with replacement, 65276 times
-    "sex" = sample(
+    "gender" = sample(
       x = c("Male", "Female", "Unidentified"),
       size = 65276,
       replace = TRUE
@@ -43,23 +43,40 @@ simulated_data <-
       size = 65276,
       replace = TRUE
     )
-)
+  )
 
-simulated_data
-
-#add a new items_found column to those events with strip search, with replacement, 32717 times (32717 is based on the set.seed() number)
+#add four new columns items_found, reason_injury, reason_escape, reason_weapons and reason_possess_evidence to those events with strip search, with replacement, 32717 times (32717 is based on the set.seed() number)
 
 set.seed(56)
 
-simulated_data_items <-
+simulated_data_items_reason <-
   simulated_data |>
   filter(strip_search == "Yes") |>
   mutate(
     "items_found" = sample(
       x = c("Yes", "No"),
       size = 32717,
-      replace = TRUE
-    )
+      replace = TRUE),
+      #randomly pick if the event includes strip search the reason is due to the person causing injury, with replacement, 32717 times
+      "reason_injury" = sample(
+        x = c(1, 0),
+        size = 32717,
+        replace = TRUE),
+      #randomly pick if the event includes strip search the reason is due to the person assisting escape, with replacement, 32717 times
+      "reason_escape" = sample(
+        x = c(1, 0),
+        size = 32717,
+        replace = TRUE),
+      #randomly pick if the event includes strip search the reason is due to the person having weapons, with replacement, 32717 times
+      "reason_weapons" = sample(
+        x = c(1, 0),
+        size = 32717,
+        replace = TRUE),
+      #randomly pick if the event includes strip search the reason is due to the person possessing evidence, with replacement, 32717 times
[…]

### 2024-01-25 outputs/paper/paper.qmd
polish charts and graphs

@@
   kable(
-    col.names = c("Number of Reasons", "Items_Found", "Count"),
+    col.names = c("Number of Reasons", "Items Found", "Count"),
     booktabs = TRUE
@@
   rename(count = n) |>
-  ggplot(aes(x = total_reasons, y = count, colour = race)) +
-  geom_point() +
+  ggplot(aes(x = total_reasons, y = count, colour = race, size = count)) +
+  geom_point(alpha = 0.5) +
   theme_bw() +
@@
 cleaned_search_reasons_data |>
-  group_by(total_reasons, items_found, race) |>
+  group_by(total_reasons, items_found, gender) |>
   count() |>
   rename(count = n) |>
-  ggplot(aes(x = total_reasons, y = count, colour = race)) +
-  geom_point() +
+  ggplot(aes(x = total_reasons, y = count, colour = gender, size = count)) +
+  geom_point(alpha = 0.5) +
   theme_bw() +
@@
        y = "Count",
-       color = "Race"
+       color = "Gender"
        ) +
   scale_fill_viridis_d(option="cividis")
+
+cleaned_search_reasons_data |>
+  group_by(total_reasons, items_found, gender) |>
+  count() |>
+  rename(count = n) |>
+  ggplot(aes(x = total_reasons, y = count, colour = items_found)) +
+  geom_bar() +
+  theme_bw() +
+  theme(legend.key.size = unit(0.5, 'cm'))+
+  labs(x = "Total Reasons for Strip Searching",
+       y = "Items Found",
+       color = "Gender"
+       ) +
+  scale_fill_viridis_d(option="cividis")
+
+cleaned_race_gender_data |>
+  ggplot(aes(x = income, y = count, label = ward_code)) + 
+  geom_point() + 
+  geom_beeswarm() +
+  geom_smooth(method=lm, se=FALSE) + 
+  geom_text(hjust=-0.5, vjust=0.5) + 
+  labs(
+    title = "Number of clinics vs. average household income for each ward",
+    x = "Average household income",
+    y = "Number of clinics",
+    label = "Ward Code"
+  ) +
+  theme_minimal()
+
 ```

### 2024-01-25 outputs/paper/paper.qmd
update graphs and bubble charts

--- 2024-01-24 outputs/paper/paper.qmd
@@
 Foremost, this paper aims to be a form of data activism to counter the injustices related to police power abuse around strip searching of the vulnerable Toronto population.
-To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using bar graphs and various other graphs.
-Particularly, for each category of race and gender, the distribution bars would be sectioned into 'including strip searches' or 'not including' to visualize the statistics of strip searching.
-With that, It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd].
+To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using stacked summary tables, stacked bar charts and bubble charts.
+Particularly, for each category of race and gender, the distribution bars or points would be sectioned into various categories like 'including strip searches' or 'not including' to visualize the statistics of strip searching for intersecting social identities With that, It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd].
 Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021.
@@
 #referenced kable stying from https://github.com/EthanSansom/torontocitytrees/blob/main/outputs/paper/paper.Rmd
-head(cleaned_race_gender_data, 10) |>
+head(cleaned_race_gender_data, 6) |>
   kable(
@@
 Vice versa, if it contains `1`, this means strip search happens because of that reason.
-In addition, a new column called "Total Reasons" totals all reasons used to account for each police strip search is added with `mutate()`, where all instances of `1` and `0` are changed into numeric with `as.numeric()`.
+In addition, a new column called "Total Reasons" totals all reasons used to account for each police strip search is added with `mutate()`, where all instances of `1` and `0` are changed into numeric with `as.numeric()` and then summed together.
 
@@
 
-  head(cleaned_search_reasons_data, 10) |>
+  head(cleaned_search_reasons_data, 6) |>
   select(-c(race, gender)) |>
@@
 #| message: false
-#| label: fig-scatter-plot-reasons
-#| fig-cap: Relationship Between Reasons for and Items Found During Per Strip Search in Toronto, 2020-2021
+#| label: fig-reasons-race
+#| fig-cap: Relationship Between Reasons for Strip Search and Race in Toronto, 2020-2021
 
-#create a scatterplot to show each observation in the dataset, comparing the number of reasons to the items found in each strip search
-#colour the points based on the police division where the group of shootings occurred
+#create a scatterplot to show each observation in the dataset, comparing the total reasons to racial identitiy found in each strip search
+#colour the points based on the perceived race of the captive
 #code for geom_point() referenced from https://github.com/TDonofrio62/Shooting-Occurrences-in-Toronto/blob/main/Shooting_Occurrences/outputs/paper/Shooting_Occurrences.Rmd
@@
   rename(count = n) |>
-  ggplot(aes(x = total_reasons, y = count, colour = race)) +
-  geom_point() +
+  ggplot(aes(x = total_reasons, y = count, colour = race, size = count)) +
+  geom_point(alpha = 0.5) +
   theme_bw() +
-  theme(legend.key.size = unit(0.5, 'cm'))+
+  theme(legend.key.size = unit(0.5, 'cm'))
   labs(x = "Total Reasons for Strip Searching",
@@
 
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| label: fig-reasons-gender
+#| fig-cap: Relationship Between Reasons for Strip Search and Gender in Toronto, 2020-2021
+
 cleaned_search_reasons_data |>
-  group_by(total_reasons, items_found, race) |>
+  group_by(total_reasons, items_found, gender) |>
   count() |>
   rename(count = n) |>
-  ggplot(aes(x = total_reasons, y = count, colour = race)) +
-  geom_point() +
+  ggplot(aes(x = total_reasons, y = count, colour = gender, size = count)) +
+  geom_point(alpha = 0.5) +
   theme_bw() +
-  theme(legend.key.size = unit(0.5, 'cm'))+
+  theme(legend.key.size = unit(0.5, 'cm'))
   labs(x = "Total Reasons for Strip Searching",
        y = "Count",
-       color = "Race"
+       color = "Gender"
        ) +
   scale_fill_viridis_d(option="cividis")
+
 ```
-
 ## Discussion

### 2024-01-25 outputs/paper/paper.qmd
refine graphs

@@
 cleaned_search_reasons_data |>
-  group_by(total_reasons, items_found, race) |>
+  group_by(total_reasons, race) |>
   count() |>
@@
 #| message: false
+#| results: hide
 #| label: fig-reasons-gender
@@
 cleaned_search_reasons_data |>
-  group_by(total_reasons, items_found, gender) |>
+  group_by(total_reasons, gender) |>
   count() |>

### 2024-01-25 outputs/paper/paper.qmd
updated charts

@@
 This paper thus inspects gender and racial identities of all Toronto police arrests from 2020 to 2021, and assesses if there is a correlation between marginalized identities and police strip searching.
-Foremost, this paper aims to be a form of data activism to counter the injustices related to police power abuse around strip searching of the vulnerable Toronto population.
+Foremost, this paper aims to be a form of data activism, contributing to the dialogue around police power abuse, using strip-search to undermine the the vulnerable populations in Toronto.
 To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using stacked summary tables, stacked bar charts and bubble charts.
@@
 This chosen dataset was published by Toronto Police Services to convey information related to all police arrests and strip searches from 2020 to 2021 (last updated on December 2, 2022).
-Originally, the dataset features information on date of arrest, event id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, the person's action at the time of arrest, and the police's reason for strip searching along with details on any items found.
+Originally, the dataset features information on date of arrest, event id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, the person's action at the time of arrest, and the police's reason for strip searching along with any items found.
+Data of strip searches are lodged based on each police arrest throughout 2020 and 2021, using a unique identifier called `event_id` which identifies the nature of the strip search and the social identities of the people involved.
+
 Informative as it is, the dataset contains much naming inconsistency and many columns outside the scope of this paper.
@@
 
-To streamline the data regarding racial and gender identities of the people involved in police strip searches, the names within the original dataset are simplified and then the dataset is split into a smaller one containing relevant columns ([see @tbl-cleaned_race_gender_data]).
-This dataset contains demographic and strip search-relevant information for each police arrest event, such as race, gender, whether the event involves a strip search, and if it is a strip search, are there any items found.
+Most relevant to the paper are the variables on the racial and gender identities of people who were arrested and then strip-searched.
+To streamline this data, the names within the original dataset are simplified and then the dataset is split into a smaller dataset containing relevant columns ([see @tbl-cleaned_race_gender_data]).
+This dataset entails demographic and strip search-relevant information for each police arrest event, such as race, gender, whether the event involves a strip search, and if it is a strip search, are there any items found.
 
@@
 This is because the items found within a strip search is only available when a police arrest does involve strip searching.
-
-Furthermore, the reason for keeping rows of events that don't include strip search is to later produce summary statistics of events with and without strip searches.
-This further aligns with the task of the paper at hand, which is to critically reflect on the rate of strip searching on people with racialized and gendered identities.
 
@@
 
+Furthermore, the reason for keeping rows of events that don't include strip search is to later produce summary statistics of events with and without strip searches.
+This further aligns with the task of the paper at hand, which is to critically reflect on the rate of strip searching on people with racialized and gendered identities.
+
 cite toronto population to say that while the number of white poeple seem large, it's not really
@@
   geom_bar(stat="identity") +
-  theme_minimal() +
   theme(legend.key.size = unit(0.5, 'cm'), axis.text.x = element_text(angle = 70, vjust = 1, hjust= 1)) +
@@
        y = "Number of Arrests",
-       fill = "Strip Search"
-  ) +
-  scale_fill_viridis_d(option="cividis")
+       fill = "Strip Search") + scale_fill_viridis_d(option="cividis")
 
@@
   ggplot(mapping = aes(x = strip_search, fill = strip_search), data=cleaned_race_gender_data) +
-  geom_bar()+theme_minimal()+ylab("number of arrests")+facet_wrap(~race)+ scale_fill_viridis_d(option="cividis")
+  geom_bar() + 
+  labs(y = "Number of Arrests", x = "Strip Search", fill = "Strip Search") +
+  facet_wrap(~race) + scale_fill_viridis_d(option="cividis")
 
@@
   geom_bar(stat="identity") + 
-  theme_minimal() +
+  coord_flip() +
   theme(legend.key.size = unit(0.5, 'cm')) +
@@
        fill = "Strip Search"
-  ) +
-  scale_fill_viridis_d(option="cividis")
+  ) + scale_fill_viridis_d(option="cividis")
 
@@
 #| message: false
+#| results: hide
 #| label: tbl-reasons-strip-search
@@
   rename(count = n) |>
-  ggplot(aes(x = total_reasons, y = count, colour = race, size = count)) +
-  geom_point(alpha = 0.5) +
-  theme_bw() +
-  theme(legend.key.size = unit(0.5, 'cm'))
+  ggplot(aes(x = total_reasons, y = count, colour = race)) +
+  geom_line(alpha = 0.5) +
+  geom_point(aes(size = count), alpha = 0.5) +
   labs(x = "Total Reasons for Strip Searching",
@@
        ) +
+  theme(legend.key.size = unit(0.5, 'cm')) +
   scale_fill_viridis_d(option="cividis")
@@
   rename(count = n) |>
-  ggplot(aes(x = total_reasons, y = count, colour = gender, size = count)) +
-  geom_point(alpha = 0.5) +
-  theme_bw() +
-  theme(legend.key.size = unit(0.5, 'cm'))
-  labs(x = "Total Reasons for Strip Searching",
-       y = "Count",
-       color = "Gender"
-       ) +
+  ggplot(aes(x = total_reasons, y = count, colour = gender)) +
+  geom_line(alpha = 0.5) +
+  geom_point(aes(size = count), alpha = 0.5) +
+  labs(y = "Count", x = "Total Reasons for Strip Searching", color = "Gender") +
+  theme(legend.key.size = unit(0.5, 'cm')) +
   scale_fill_viridis_d(option="cividis")
@@
 ```
+
 ## Discussion

### 2024-01-25 outputs/paper/paper.qmd
update results

@@
 
-The other dataset conveys all police arrests that involve strip searching and the reasons behind strip searching.
+To further inspect the nature of each strip search, reasons for strip searching prove a necessary data source.
+Thus, this other dataset conveys all police arrests that involve strip searching and the reasons behind strip searching.
 Under close examination, the naming of individual data cells in different columns from the original, raw dataset is quite disordered, for instance, using `None` and `XX` in place of the usual `NA`.
@@
 
-Furthermore, the reason for keeping rows of events that don't include strip search is to later produce summary statistics of events with and without strip searches.
-This further aligns with the task of the paper at hand, which is to critically reflect on the rate of strip searching on people with racialized and gendered identities.
+Throughout 2020 and 2021, there are a total of 7801 recorded strip searches out of 65276 incidents of police arrests, which accounts for 12% of all arrests [see @tbl-reasons-strip-search].
+Excluding those whose racial identities are unknown and non-existent, by percentage, the three most affected racial groups are, in descending order, of Indigenous, African and Caucasian descent, standing at 16%, 14% and 13% each [see @tbl-race-strip-search].
+Gender-wise, men are more likely to be strip searched compared to women, at 12% and 10% respectively (@tbl-gender-strip-search).
+Though, observably, the difference between the two sexes quite marginal.
+People of unidentified gender identity aren't strip searched at all according to the dataset (denoted as `U` in the tables and graphs).
+
+As of January 22, 2023, there are 434 COVID-19 immunization clinics across the City of Toronto.
+Further breakdown shows that there are 5 clinics that are city operated, 16 clinics that are run by hospitals, and 415 are offered in local pharmacies.
+On average, there are 17 clinics per ward in Toronto, with standard deviation of 6.
+The wards with the highest number of clinics are: Toronto Center (Ward 13) with 29 clinics, Etobicoke-Lakeshore (Ward 3) with 28 clinics, and University-Rosedale (Ward 11) with 28 clinics.
+The wards with the lowest number of clinics are: Humber River-Black Creek (Ward 7) with 9 clinics, Davenport (Ward 9) with 12 clinics, Scarborough Southwest (Ward 20) with 12 clinics, Scarborough-Agincourt (Ward 22) with 12 clinics, and Scarborough-Rouge Park (Ward 25) with 12 clinics.
+
+In add
 
@@
 ### Racialized and Gendered Police Strip Searches Statistics
+
+In addition, a new column called "Total Reasons" totals all reasons used to account for each police strip search is added with `mutate()`, where all instances of `1` and `0` are changed into numeric with `as.numeric()` and then summed together.
+to ensure that all data are comparative to one another
 
@@
 
-```{r fig.height=6, fig.width=7}
+```{r fig.width=8, fig.height=5}
 #| echo: false
@@
   rename(number_of_arrests = n) |>
-  ggplot(aes(x = race, y = number_of_arrests, fill = strip_search)) + 
+  group_by(race) |>
+  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
+  ggplot(aes(x = race, y = percentage, fill = strip_search)) + 
   geom_bar(stat="identity") +
-  theme(legend.key.size = unit(0.5, 'cm'), axis.text.x = element_text(angle = 70, vjust = 1, hjust= 1)) +
+  coord_flip() +
+  theme_minimal() +
+  theme(legend.key.size = unit(0.5, 'cm')) +
   labs(x = "Race",
-       y = "Number of Arrests",
+       y = "Percentage",
        fill = "Strip Search") + scale_fill_viridis_d(option="cividis")
@@
   rename(number_of_arrests = n) |>
-  filter(strip_search == "Yes") |> 
-  ggplot(mapping = aes(x = strip_search, fill = strip_search), data=cleaned_race_gender_data) +
+  group_by(race) |>
+  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
+  ggplot(mapping = aes(x = gender, fill = strip_search), data=cleaned_race_gender_data) +
   geom_bar() + 
-  labs(y = "Number of Arrests", x = "Strip Search", fill = "Strip Search") +
-  facet_wrap(~race) + scale_fill_viridis_d(option="cividis")
+  labs(y = "Percentage", x = "Strip Search", fill = "Strip Search") +
+  facet_wrap(~race) + scale_fill_viridis_d(option="cividis") +
 
@@
   coord_flip() +
+  theme_minimal() +
   theme(legend.key.size = unit(0.5, 'cm')) +
@@
   ) + scale_fill_viridis_d(option="cividis")
+
+cleaned_race_gender_data |>
+  group_by(gender, strip_search) |>
+  count() |>
+  rename(number_of_arrests = n) |>
+  group_by(gender) |>
+  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
+  ggplot(aes(x = gender, y = percentage, fill = strip_search)) + 
+  geom_bar(stat="identity") + 
+  coord_flip() +
+  theme_minimal() +
+  theme(legend.key.size = unit(0.5, 'cm')) +
+  labs(x = "Gender",
+       y = "Percentage",
+       fill = "Strip Search"
+  ) + scale_fill_viridis_d(option="cividis")
+
+cleaned_race_gender_data |>
+  group_by(gender, strip_search) |>
+  count() |>
+  rename(number_of_arrests = n) |>
+  group_by(gender) |>
+  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
+  ggplot(mapping = aes(x = strip_search, fill = strip_search), data=cleaned_race_gender_data) +
+  geom_bar() + 
+  labs(y = "Number of Arressts", x = "Strip Search", fill = "Strip Search") +
+  facet_wrap(~gender) + scale_fill_viridis_d(option="cividis")
 
@@
        ) +
+  theme_minimal() +
   theme(legend.key.size = unit(0.5, 'cm')) +
@@
   theme(legend.key.size = unit(0.5, 'cm')) +
+  theme_minimal() +
   scale_fill_viridis_d(option="cividis")
@@
 
+binaristic view of gender
+
 ## Conclusion

### 2024-01-25 outputs/paper/paper.qmd
update tables, graphs and results

--- 2024-01-25 outputs/paper/paper.qmd
@@
 bibliography: references.bib
+output: 
+  pdf_document :
+    keep_tex: true
+    number_sections: true
+header-includes:
+ \usepackage{wrapfig}
+ \usepackage{float}
+ \floatplacement{figure}{H}
 ---
@@
 library(here)
-library(ggbeeswarm)
 library(ggrepel)
 library(ggplot2)
+library(float)
+
+citation("float")
 
@@
 Foremost, this paper aims to be a form of data activism, contributing to the dialogue around police power abuse, using strip-search to undermine the the vulnerable populations in Toronto.
-To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using stacked summary tables, stacked bar charts and bubble charts.
+To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using stacked summary tables, stacked histograms and bubble charts.
 Particularly, for each category of race and gender, the distribution bars or points would be sectioned into various categories like 'including strip searches' or 'not including' to visualize the statistics of strip searching for intersecting social identities With that, It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd].
@@
 
-This chosen dataset was published by Toronto Police Services to convey information related to all police arrests and strip searches from 2020 to 2021 (last updated on December 2, 2022).
-Originally, the dataset features information on date of arrest, event id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, the person's action at the time of arrest, and the police's reason for strip searching along with any items found.
+Published by Toronto Police Services, this dataset conveys information related to all police arrests and strip searches from 2020 to 2021 (last updated on December 2, 2022).
+Originally, the dataset features 26 varibles, including date of arrest, event id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, the person's action at the time of arrest, and the police's reason for strip searching along with any items found.
 Data of strip searches are lodged based on each police arrest throughout 2020 and 2021, using a unique identifier called `event_id` which identifies the nature of the strip search and the social identities of the people involved.
 
-Informative as it is, the dataset contains much naming inconsistency and many columns outside the scope of this paper.
+Informative as it is, the dataset contains much naming inconsistency and many variables outside the scope of this paper.
 Thus, in order to conduct an intersectional---or gender and race---analysis of police strip searches, I have to further clean this dataset and split it into two smaller datasets.
@@
 
-The other dataset conveys all police arrests that involve strip searching and the reasons behind strip searching.
-Under close examination, the naming of individual data cells in different columns from the original, raw dataset is quite disordered, for instance, using `None` and `XX` in place of the usual `NA`.
+To further inspect the nature of each strip search, reasons for strip searching prove a vital data source.
+Thus, this dataset--another sub-dataset split from the orginal raw data--conveys all police arrests that involve strip searching and the reasons behind strip searching.
+Under close examination, the naming of individual data cells from the original, raw dataset is quite disordered, for instance, using `None` and `XX` in place of the usual `NA`.
 Thus, all inconsistent data cells were renamed to streamline the naming conventions and data processing later on.
-After renaming, the big dataset was split into containing relevant columns ([see @tbl-cleaned_search_reasons_data]).
-Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for rows with `NA` as values for all the `reason` column.
+After renaming, the big dataset was refined into containing relevant columns ([see @tbl-cleaned_search_reasons_data]).
+Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for rows with `NA` as values for all of the `reason` variables
 
-Except for "ID" column, all other columns in this new dataset are reasons which account for why the Toronto police force conducts strip search under certain arrests.
+Except for "ID" variable, all other variables in this new dataset are reasons to account for why the Toronto police force conducts strip search under certain arrests.
 For instance, "Reason - Assist Escape" is for people who may have helped other suspects escape and "Reason - Possess Evidence" is for people who are suspected of harboring items like drugs or an alibi.
-For each "Reason" column, if the data cell contains `0`, this means that strip search doesn't happen because of that particular reason.
-Vice versa, if it contains `1`, this means strip search happens because of that reason.
-In addition, a new column called "Total Reasons" totals all reasons used to account for each police strip search is added with `mutate()`, where all instances of `1` and `0` are changed into numeric with `as.numeric()` and then summed together.
+For each "Reason" variable, if the data cell contains `0`, this means that strip search doesn't happen.
+Vice versa, if it contains `1`, this means strip search happens.
+In addition, a new column called "Total Reasons" is added to sum  all reasons used to account for each strip search.
+To do this, `mutate()` is utilized, where all instances of `1` and `0` are changed into numeric with `as.numeric()` and then summed together.
 
@@
 
-Furthermore, the reason for keeping rows of events that don't include strip search is to later produce summary statistics of events with and without strip searches.
-This further aligns with the task of the paper at hand, which is to critically reflect on the rate of strip searching on people with racialized and gendered identities.
+### Racialized and Gendered Police Strip Searches Statistics
+
+To encapsulate, throughout 2020 and 2021, there are a total of 7801 recorded strip searches out of 65276 incidents of police arrests, which accounts for a whooping 12% of all arrests (see [@tbl-items-strip-search]).
+Excluding those whose racial identities are unknown and non-existent, by percentage, the three most affected racial groups are, in descending order, of Indigenous, Black and White ethnic roots, standing at 16%, 14% and 13% each (see [@tbl-race-strip-search]).
+Gender-wise, men are more likely to be strip searched compared to women, at 12% and 10% respectively (see [@tbl-gender-strip-search]).
+Though, observably, the difference is quite marginal.
+To note, people of unidentified gender identity aren't strip searched at all according to the dataset (denoted as `U` in the tables and graphs).
+
+It's worth noting that to account for the discrepancies between counts of strip search grouped by different racial identities are challenging.
+This is due to the incongruous population spread of each racial group within Toronto.
+For instance, Toronto is largely made up of people from Caucasian descent, thus it's more probable that the people arrested and strip searched by the police force are largely white in the dataset (see [@@tbl-race-strip-search]) Therefore, to render the data statistically proportionate to one another,
 
 cite toronto population to say that while the number of white poeple seem large, it's not really
-
-### Racialized and Gendered Police Strip Searches Statistics
 
@@
   count() |>
-  rename(number_of_arrests = n) |>
+  rename(count = n) |>
   group_by(race) |>
-  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
+  mutate(percentage = round(count/sum(count)*100),
+         total_arrests = sum(count)) |>
   kable(
-    col.names = c("Race", "Strip Search", "Number of Arrests", "Percentage"),
+    col.names = c("Race", "Strip Search", "Count", "Percentage", "Total Arrests"),
     booktabs = TRUE)
@@
 
-```{r fig.height=6, fig.width=7}
+In addition, a new column called "Total Reasons" totals all reasons used to account for each police strip search is added with `mutate()`, where all instances of `1` and `0` are changed into numeric with `as.numeric()` and then summed together.
+to ensure that all data are comparative to one another
+
+```{r}
 #| echo: false
@@
   count() |>
-  rename(number_of_arrests = n) |>
-  ggplot(aes(x = race, y = number_of_arrests, fill = strip_search)) + 
+  rename(count = n) |>
+  group_by(race) |>
+  mutate(percentage = round(count/sum(count)*100)) |>
+  ggplot(aes(x = race, y = percentage, fill = strip_search)) + 
   geom_bar(stat="identity") +
-  theme(legend.key.size = unit(0.5, 'cm'), axis.text.x = element_text(angle = 70, vjust = 1, hjust= 1)) +
+  coord_flip() +
+  theme_minimal() +
+  theme(legend.key.size = unit(0.5, 'cm')) +
   labs(x = "Race",
-       y = "Number of Arrests",
+       y = "Percentage",
        fill = "Strip Search") + scale_fill_viridis_d(option="cividis")
@@
   count() |>
-  rename(number_of_arrests = n) |>
-  filter(strip_search == "Yes") |> 
-  ggplot(mapping = aes(x = strip_search, fill = strip_search), data=cleaned_race_gender_data) +
+  rename(count= n) |>
+  group_by(race) |>
+  mutate(percentage = round(count/sum(count)*100)) |>
+  ggplot(mapping = aes(x = gender, fill = strip_search), data=cleaned_race_gender_data) +
   geom_bar() + 
-  labs(y = "Number of Arrests", x = "Strip Search", fill = "Strip Search") +
+  labs(y = "Percentage", x = "Strip Search", fill = "Strip Search") +
   facet_wrap(~race) + scale_fill_viridis_d(option="cividis")
@@
   count() |>
-  rename(number_of_arrests = n) |>
+  rename(count = n) |>
   group_by(gender) |>
-  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
+  mutate(percentage = round(count/sum(count)*100),
+         total_arrests = sum(count)) |>
   kable(
-    col.names = c("Gender", "Strip Search", "Number of Arrests", "Percentage"),
+    col.names = c("Gender", "Strip Search", "Count", "Percentage", "Total Arrests"),
     booktabs = TRUE)
@@
   count() |>
-  rename(number_of_arrests = n) |>
-  ggplot(aes(x = gender, y = number_of_arrests, fill = strip_search)) + 
+  rename(count = n) |>
+  ggplot(aes(x = gender, y = count, fill = strip_search)) + 
   geom_bar(stat="identity") + 
   coord_flip() +
+  theme_minimal() +
   theme(legend.key.size = unit(0.5, 'cm')) +
   labs(x = "Gender",
-       y = "Number of Arrests",
+       y = "Count",
        fill = "Strip Search"
   ) + scale_fill_viridis_d(option="cividis")
+
+cleaned_race_gender_data |>
+  group_by(gender, strip_search) |>
+  count() |>
+  rename(count = n) |>
+  group_by(gender) |>
+  mutate(percentage = round(count/sum(count)*100)) |>
+  ggplot(aes(x = gender, y = percentage, fill = strip_search)) + 
+  geom_bar(stat="identity") + 
+  coord_flip() +
+  theme_minimal() +
+  theme(legend.key.size = unit(0.5, 'cm')) +
+  labs(x = "Gender",
+       y = "Percentage",
+       fill = "Strip Search"
+  ) + scale_fill_viridis_d(option="cividis")
+
+cleaned_race_gender_data |>
+  group_by(gender, strip_search) |>
+  count() |>
+  rename(count = n) |>
+  group_by(gender) |>
+  mutate(percentage = round(count/sum(count)*100)) |>
+  ggplot(mapping = aes(x = strip_search, fill = strip_search), data=cleaned_race_gender_data) +
+  geom_bar() + 
+  labs(y = "Count", x = "Strip Search", fill = "Strip Search") +
+  facet_wrap(~gender) + scale_fill_viridis_d(option="cividis")
 
@@
 
+As for reasons of strip searching provided by the Toronto Police Services and items found along these strip searches, there is a substantial 10% of strip searches being conducted without any reason (see [@tbl-reasons-strip-search]).
+Most commonly, strip searches are carried out under at least 1 or 2 reasons, both at 29% (see [@tbl-reasons-strip-search]).
+Overall, most of strip search enforcement doesn't garner any items in question from the people who were suspected by the police.
+Specifically, a whooping 63% of strip searches don't have any items found (see [@tbl-items-strip-search]).
+
 ```{r}
@@
 #| message: false
-#| results: hide
-#| label: tbl-reasons-strip-search
+#| label: tbl-items-strip-search
 #| tbl-cap: Summary Statistics of Items Found During Police Strip Searches
@@
   count() |>
-  rename(number_of_arrests = n) |>
+  rename(count = n) |>
   filter(strip_search == "Yes") |>
-  select(items_found, number_of_arrests) |>
-  kable(
-    col.names = c("Strip Search", "Items Found", "Number of Arrests"),
-    booktabs = TRUE
+  select(items_found, count) |>
+  group_by(strip_search) |>
+  mutate(total_searches = sum(count),
+         percentage_of_items_found = round(count/total_searches*100)) |>
+  kable(col.names = c("Strip Search", "Items Found", "Count", "Total Searches", "Percentage of Items Found"), booktabs = TRUE
         )
@@
 #| message: false
-#| label: tbl-items-strip-search
+#| label: tbl-reasons-strip-search
 #| tbl-cap: Summary Statistics of Reasons for Police Strip Searches
@@
   rename(count = n) |>
+  group_by(total_reasons) |>
+  mutate(percentage = round(count/sum(count)*100),
+         total_searches = sum(count),
+         percentage_total = round(total_searches/7801*100)) |>
   kable(
-    col.names = c("Number of Reasons", "Items Found", "Count"),
+    col.names = c("Total Reasons", "Items Found", "Count", "Percentage by Reasons", "Total Searches", "Percentage by Searches"),
     booktabs = TRUE
-  )
+  ) |> 
+  kable_styling(latex_options="scale_down")
 ```
@@
 
+binaristic view of gender
+
+mystified, how is race sectioned and divided
+
+biased source, observation
+
 ## Conclusion

### 2024-01-25 outputs/paper/paper.qmd
new references
clean formatting of graphs

@@
 ---
-title: "Slight Positive Correlation Between Gendered and Racialized Police Strip Searches in the Toronto Municipal, 2020-2021"
+title: "Slightly Positive Correlation Between Racialization, Gendering and Police Strip Searches in the Toronto Municipal, 2020-2021"
 title-block-banner: true
 abstract: "Police strip search has been a topic of great contention during the COVID-19 pandemic.
-The question commonly raised was how certain demographics are more likely to undergo arrests and strip searches than others, and why has this skewed power dynamics been maintained for so long.
+The question commonly raised was how certain demographics are more likely to undergo strip searches than others, and why has this skewed power dynamics been maintained for so long.
 Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021.
-This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities.
-More on RESULTS"
+In short, this paper finds that there is a slight positive correlation between police strip searches and the gendered and racialized Toronto demographic such as females or people of African descent.
+This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities."
 thanks: "Code and data are available at: https://github.com/ponolite/Police_Arrests_and_Strip_Searches/tree/main"
@@
 bibliography: references.bib
-output: 
-  pdf_document :
-    keep_tex: true
-    number_sections: true
-header-includes:
- \usepackage{wrapfig}
- \usepackage{float}
- \floatplacement{figure}{H}
 ---
@@
 library(float)
-
-citation("float")
 
@@
 This paper thus inspects gender and racial identities of all Toronto police arrests from 2020 to 2021, and assesses if there is a correlation between marginalized identities and police strip searching.
-Foremost, this paper aims to be a form of data activism, contributing to the dialogue around police power abuse, using strip-search to undermine the the vulnerable populations in Toronto.
-To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using stacked summary tables, stacked histograms and bubble charts.
-Particularly, for each category of race and gender, the distribution bars or points would be sectioned into various categories like 'including strip searches' or 'not including' to visualize the statistics of strip searching for intersecting social identities With that, It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd].
+Foremost, this paper aims to be a form of data activism, contributing to the dialogue around police power abuse that enforces strip-search to undermine the the vulnerable populations in Toronto.
+To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using summary tables, stacked histograms and bubble charts.
+Particularly, for each category of race and gender, the distribution bars or points would be sectioned into various categories like 'including strip searches' or 'not including' to visualize the statistics of strip searching against intersecting social identities.
+It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd].
 Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021.
-Yet, it's still worth looking at due the precedents of police power abuse [@covidStripSearch].
+Nonetheless, it's still worth looking at due the precedents of police power abuse [@covidStripSearch].
 
-Structurally then, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
+This paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
 First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data cleaning and extraction took place [@torontoPolice].
@@
 Only one dataset was retrieved from the data package to examine the racial and gender identities of Toronto police arrests from 2020 to 2021, which is the package `Police Race and Identity Based Data - Arrests and Strip Searches` [@torontoPolice].
-Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra], `naniar` [@rNaniar] and `knitr` [@rKnitr].
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `here` [@rHere], `ggrepel` [@rGgrepel], janitor` [@rJanitor], `kableExtra` [@rKableExtra], `naniar` [@rNaniar] and `knitr` [@rKnitr].
 
@@
 
-Published by Toronto Police Services, this dataset conveys information related to all police arrests and strip searches from 2020 to 2021 (last updated on December 2, 2022).
-Originally, the dataset features 26 varibles, including date of arrest, event id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, the person's action at the time of arrest, and the police's reason for strip searching along with any items found.
+Published by Toronto Police Services, this primary dataset conveys information related to all police arrests and strip searches from 2020 to 2021 (which halted refreshing on December 2, 2022).
+Originally, the dataset features 26 variables, including date of arrest, event id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, the person's action at the time of arrest, and the police's reason for strip searching along with any items found.
 Data of strip searches are lodged based on each police arrest throughout 2020 and 2021, using a unique identifier called `event_id` which identifies the nature of the strip search and the social identities of the people involved.
+No other dataset was considered for the paper but this one, since it's the only dataset that communicates information on police enforcement of strip searching.
 
@@
     col.names = c("ID", "Race", "Gender", "Strip Search", "Items Found"),
-    booktabs = TRUE
+    booktabs = TRUE,
+    align = c("c", "c", "c", "c", "c")
   )
@@
 Thus, this dataset--another sub-dataset split from the orginal raw data--conveys all police arrests that involve strip searching and the reasons behind strip searching.
-Under close examination, the naming of individual data cells from the original, raw dataset is quite disordered, for instance, using `None` and `XX` in place of the usual `NA`.
+Under close examination, the naming of individual data cells from the original, raw dataset is quite disordered, for instance, using `None` and `XX` in place of the usual `NA` to differentiate between variables.
 Thus, all inconsistent data cells were renamed to streamline the naming conventions and data processing later on.
 After renaming, the big dataset was refined into containing relevant columns ([see @tbl-cleaned_search_reasons_data]).
-Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for rows with `NA` as values for all of the `reason` variables
+Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for all `reason` variables with `NA` as their values.
 
-Except for "ID" variable, all other variables in this new dataset are reasons to account for why the Toronto police force conducts strip search under certain arrests.
+Except for "ID" variable, the four other variables in this new dataset are reasons to account for why the Toronto police force conducts strip search under certain arrests ([see @tbl-cleaned_search_reasons_data]).
 For instance, "Reason - Assist Escape" is for people who may have helped other suspects escape and "Reason - Possess Evidence" is for people who are suspected of harboring items like drugs or an alibi.
-For each "Reason" variable, if the data cell contains `0`, this means that strip search doesn't happen.
-Vice versa, if it contains `1`, this means strip search happens.
+For each "Reason" variable, if the data cell contains `0`, this means that strip search happens for other reasons beside this particular one.
+Vice versa, if it contains `1`, this means strip search happens because of this very reason.
 In addition, a new column called "Total Reasons" is added to sum  all reasons used to account for each strip search.
 To do this, `mutate()` is utilized, where all instances of `1` and `0` are changed into numeric with `as.numeric()` and then summed together.
+
+\newpage
 
@@
     col.names = c("ID", "Reason - Injury", "Reason - Escape", "Reason - Weapons", "Reason - Has Evidence", "Items_Found", "Total Reasons"),
-    booktabs = TRUE
+    booktabs = TRUE,
+    align = c("c", "c", "c", "c", "c", "c", "c")
   ) |>
@@
 To note, people of unidentified gender identity aren't strip searched at all according to the dataset (denoted as `U` in the tables and graphs).
-
-It's worth noting that to account for the discrepancies between counts of strip search grouped by different racial identities are challenging.
-This is due to the incongruous population spread of each racial group within Toronto.
-For instance, Toronto is largely made up of people from Caucasian descent, thus it's more probable that the people arrested and strip searched by the police force are largely white in the dataset (see [@@tbl-race-strip-search]) Therefore, to render the data statistically proportionate to one another,
-
-cite toronto population to say that while the number of white poeple seem large, it's not really
 
@@
     col.names = c("Race", "Strip Search", "Count", "Percentage", "Total Arrests"),
-    booktabs = TRUE)
+    booktabs = TRUE,
+    align = c("c", "c", "c", "c", "c"))
 
 ```
+
+Particularly, accounting for the discrepancies between counts of strip search grouped by different racial and gender identities are challenging.
+This is due to the incongruous population spread of each racial and gender group within Toronto.
+For instance, a major 44.3% of Toronto population is white people [@ethnicCensus], thus it's more probable that the people arrested and strip searched by the police force are largely `white` in the dataset (see [@@tbl-race-strip-search]) As such, to render the data statistically proportionate to one another, `mutate()`was largely used to create new variables that are mostly percentage values.
+Admittingly so, this enables each variable to be on common grounds when compared with one another.
 
@@
 
-```{r}
+```{r fig.width = 8}
 #| echo: false
@@
   ggplot(aes(x = race, y = percentage, fill = strip_search)) + 
-  geom_bar(stat="identity") +
-  coord_flip() +
+  geom_bar(stat="identity", position = "dodge") +
   theme_minimal() +
-  theme(legend.key.size = unit(0.5, 'cm')) +
+  theme(legend.key.size = unit(0.5, 'cm'), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
   labs(x = "Race",
@@
 
-```{r fig.height=6, fig.width=7}
+```{r fig.height=6, fig.width=8}
 #| echo: false
@@
 cleaned_race_gender_data |>
-  group_by(race, strip_search) |>
+  group_by(race, strip_search, gender) |>
   count() |>
@@
   ggplot(mapping = aes(x = gender, fill = strip_search), data=cleaned_race_gender_data) +
-  geom_bar() + 
-  labs(y = "Percentage", x = "Strip Search", fill = "Strip Search") +
+  geom_bar(position = "dodge") + 
+  labs(y = "Count", x = "Gender", fill = "Strip Search") +
   facet_wrap(~race) + scale_fill_viridis_d(option="cividis")
@@
     col.names = c("Gender", "Strip Search", "Count", "Percentage", "Total Arrests"),
-    booktabs = TRUE)
+    booktabs = TRUE,
+    align = c("c", "c", "c", "c", "c"))
 
@@
 
-```{r}
+```{r fig.height=3, fig.width=8}
 #| echo: false
@@
   rename(count = n) |>
-  ggplot(aes(x = gender, y = count, fill = strip_search)) + 
-  geom_bar(stat="identity") + 
-  coord_flip() +
-  theme_minimal() +
-  theme(legend.key.size = unit(0.5, 'cm')) +
-  labs(x = "Gender",
-       y = "Count",
-       fill = "Strip Search"
-  ) + scale_fill_viridis_d(option="cividis")
-
-cleaned_race_gender_data |>
-  group_by(gender, strip_search) |>
-  count() |>
-  rename(count = n) |>
-  group_by(gender) |>
-  mutate(percentage = round(count/sum(count)*100)) |>
-  ggplot(aes(x = gender, y = percentage, fill = strip_search)) + 
-  geom_bar(stat="identity") + 
-  coord_flip() +
-  theme_minimal() +
-  theme(legend.key.size = unit(0.5, 'cm')) +
-  labs(x = "Gender",
-       y = "Percentage",
-       fill = "Strip Search"
-  ) + scale_fill_viridis_d(option="cividis")
-
-cleaned_race_gender_data |>
-  group_by(gender, strip_search) |>
-  count() |>
-  rename(count = n) |>
   group_by(gender) |>
@@
   geom_bar() + 
+  theme_minimal() +
   labs(y = "Count", x = "Strip Search", fill = "Strip Search") +
@@
 ```
+
+\break
 
@@
          percentage_of_items_found = round(count/total_searches*100)) |>
-  kable(col.names = c("Strip Search", "Items Found", "Count", "Total Searches", "Percentage of Items Found"), booktabs = TRUE
+  kable(col.names = c("Strip Search", "Items Found", "Count", "Total Searches", "Percentage of Items Found"), booktabs = TRUE, align = c("c", "c", "c", "c", "c")
         )
@@
     col.names = c("Total Reasons", "Items Found", "Count", "Percentage by Reasons", "Total Searches", "Percentage by Searches"),
-    booktabs = TRUE
+    booktabs = TRUE,
+    align = c("c", "c", "c", "c", "c", "c")
   ) |> 
@@
 
-```{r}
+```{r fig.height=4, fig.width=8}
 #| echo: false
@@
 #| fig-cap: Relationship Between Reasons for Strip Search and Race in Toronto, 2020-2021
+#| fig-subcap: ["Race","Gender"]
+#| layout-ncol: 2
 
@@
   ggplot(aes(x = total_reasons, y = count, colour = race)) +
-  geom_line(alpha = 0.5) +
-  geom_point(aes(size = count), alpha = 0.5) +
+  geom_line() +
+  geom_point(aes(size = count)) +
   labs(x = "Total Reasons for Strip Searching",
@@
 
+cleaned_search_reasons_data |>
+  group_by(total_reasons, gender) |>
+  count() |>
+  rename(count = n) |>
+  ggplot(aes(x = total_reasons, y = count, colour = gender)) +
+  geom_line() +
+  geom_point(aes(size = count)) +
+  labs(y = "Count", x = "Total Reasons for Strip Searching", color = "Gender") +
+  theme(legend.key.size = unit(0.5, 'cm')) +
+  theme_minimal() +
+  scale_fill_viridis_d(option="cividis")
+
 ```
 
-```{r}
+```{r fig.height=3, fig.width=8}
 #| echo: false
@@
   ggplot(aes(x = total_reasons, y = count, colour = gender)) +
-  geom_line(alpha = 0.5) +
-  geom_point(aes(size = count), alpha = 0.5) +
+  geom_line() +
+  geom_point(aes(size = count)) +
   labs(y = "Count", x = "Total Reasons for Strip Searching", color = "Gender") +
@@
 
+lack tornotno ethnic demofraphic data to actually compare and contrast if the perecentage of strip search is correlated to the population spread
+
 ## Conclusion
+
+## LLMs
+
+Statement on LLM usage: no Learning Management Systems (LMSs) were solicited during the writing of this paper.
 

### 2024-01-26 outputs/paper/paper.qmd
update dicussion and graphs

@@
 Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021.
-In short, this paper finds that there is a slight positive correlation between police strip searches and the gendered and racialized Toronto demographic such as females or people of African descent.
+In short, this paper finds that police strip searches are slightly racialized and gendered, especially towards people of African descent and females within Toronto.
 This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities."
@@
 Surprisingly, while contended as a dehumanizing policing act, little has been done to assess strip search as a form of police power abuse in Toronto.
-This is especially relevant considering the city's racialized and gendered demographics, how strip searches most likely affect these social groups under the subtext of COVID-19 from 2020 to 2021 [@stripSearchOpEd].
+This is especially relevant considering the city's racialized and gendered demographics [@ethnicCensus; @genderCensus], how strip searches most likely affect these social groups under the subtext of COVID-19 from 2020 to 2021 [@stripSearchOpEd].
 
@@
 Foremost, this paper aims to be a form of data activism, contributing to the dialogue around police power abuse that enforces strip-search to undermine the the vulnerable populations in Toronto.
-To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using summary tables, stacked histograms and bubble charts.
+To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using summary tables, stacked bar charts and bubble charts.
 Particularly, for each category of race and gender, the distribution bars or points would be sectioned into various categories like 'including strip searches' or 'not including' to visualize the statistics of strip searching against intersecting social identities.
@@
 
-### Racialized and Gendered Police Strip Searches Statistics
+### Racialized Police Strip Searches Statistics
 
 To encapsulate, throughout 2020 and 2021, there are a total of 7801 recorded strip searches out of 65276 incidents of police arrests, which accounts for a whooping 12% of all arrests (see [@tbl-items-strip-search]).
-Excluding those whose racial identities are unknown and non-existent, by percentage, the three most affected racial groups are, in descending order, of Indigenous, Black and White ethnic roots, standing at 16%, 14% and 13% each (see [@tbl-race-strip-search]).
-Gender-wise, men are more likely to be strip searched compared to women, at 12% and 10% respectively (see [@tbl-gender-strip-search]).
-Though, observably, the difference is quite marginal.
-To note, people of unidentified gender identity aren't strip searched at all according to the dataset (denoted as `U` in the tables and graphs).
+Excluding those whose racial identities are unknown and non-existent, by percentage, the three most affected racial groups are, in descending order, of Indigenous, black and white background, standing at 16%, 14% and 13% each (see [@tbl-race-strip-search]).
 
@@
 This is due to the incongruous population spread of each racial and gender group within Toronto.
-For instance, a major 44.3% of Toronto population is white people [@ethnicCensus], thus it's more probable that the people arrested and strip searched by the police force are largely `white` in the dataset (see [@@tbl-race-strip-search]) As such, to render the data statistically proportionate to one another, `mutate()`was largely used to create new variables that are mostly percentage values.
-Admittingly so, this enables each variable to be on common grounds when compared with one another.
+For instance, 44.3% of Toronto population is made up of white people [@ethnicCensus], thus it's more probable that the people arrested and strip searched by the police force are ethnically "White" in the dataset (see [@tbl-race-strip-search]).
+As such, to render the data statistically proportionate to one another, `mutate()`was used to create new variables that are mostly percentage values.
 
-In addition, a new column called "Total Reasons" totals all reasons used to account for each police strip search is added with `mutate()`, where all instances of `1` and `0` are changed into numeric with `as.numeric()` and then summed together.
-to ensure that all data are comparative to one another
+Then, @fig-race-strip-search illustrates the percentage of strip searches among different racial groups in Toronto from 2020-2021.
+Observably, while being the the major population group in Toronto, people of white racial roots aren't the most strip searched.
+Rather, it's the racialized population of Toronto like black and Indigenous people that are being strip searched by police.
+This hints at the possibility of strip search being more racialized than it seems, targeting minority groups in Toronto.
 
-```{r fig.width = 8}
+```{r fig.height = 4, fig.width = 8}
 #| echo: false
 #| message: false
-#| label: fig-race_strip_search
+#| label: fig-race-strip-search
 #| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
@@
 
-```{r fig.height=6, fig.width=8}
-#| echo: false
-#| message: false
-#| label: fig-race_strip_search_separate
-#| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
+\break
 
-cleaned_race_gender_data |>
-  group_by(race, strip_search, gender) |>
-  count() |>
-  rename(count= n) |>
-  group_by(race) |>
-  mutate(percentage = round(count/sum(count)*100)) |>
-  ggplot(mapping = aes(x = gender, fill = strip_search), data=cleaned_race_gender_data) +
-  geom_bar(position = "dodge") + 
-  labs(y = "Count", x = "Gender", fill = "Strip Search") +
-  facet_wrap(~race) + scale_fill_viridis_d(option="cividis")
-
-```
+### Gendered Police Strip Searches Statistics
 
@@
 
-```{r fig.height=3, fig.width=8}
+Gender-wise, men are more likely to be strip searched compared to women, at 12% and 10% respectively (see [@tbl-gender-strip-search]).
+Though, observably, the difference is quite marginal.
+To note, people of unidentified gender identity aren't strip searched at all according to the dataset (denoted as `U` in the tables and graphs).
+
+@fig-gender-strip-search illustrates the percentage of gender groups being strip searched in Toronto.
+It's observable that the number of strip searches between men and women are comparable, especially considering the 48-52% population split between the two categories of gender in Toronto, according to the 2021 Census [@genderCensus].
+
+Nonetheless, the dataset only contains a binaristic view of gender, without regarding those who identify as non-binary or others, and essentially lumping all other gender identities as "Unidentified".
+This proves inconsistent against the constant public backlash agasint police for their power abuse of people belonging to the LGBTQI+ community in Toronto [@policeGender].
+This point will be further discussed below.
+
+```{r fig.height=3, fig.width=5}
 #| echo: false
 #| message: false
-#| label: fig-gender_strip_search
+#| label: fig-gender-strip-search
 #| fig-cap: Gender Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
@@
   mutate(percentage = round(count/sum(count)*100)) |>
-  ggplot(mapping = aes(x = strip_search, fill = strip_search), data=cleaned_race_gender_data) +
-  geom_bar() + 
+  ggplot(aes(x = gender, y = percentage, fill = strip_search)) + 
+  geom_bar(stat="identity", position = "dodge") +
   theme_minimal() +
-  labs(y = "Count", x = "Strip Search", fill = "Strip Search") +
-  facet_wrap(~gender) + scale_fill_viridis_d(option="cividis")
+  theme(legend.key.size = unit(0.5, 'cm')) +
+  labs(x = "Race",
+       y = "Percentage",
+       fill = "Strip Search") + scale_fill_viridis_d(option="cividis")
 
@@
 
-\break
-
 ### Police Strip Searches and Items Found Statistics
 
-As for reasons of strip searching provided by the Toronto Police Services and items found along these strip searches, there is a substantial 10% of strip searches being conducted without any reason (see [@tbl-reasons-strip-search]).
-Most commonly, strip searches are carried out under at least 1 or 2 reasons, both at 29% (see [@tbl-reasons-strip-search]).
-Overall, most of strip search enforcement doesn't garner any items in question from the people who were suspected by the police.
-Specifically, a whooping 63% of strip searches don't have any items found (see [@tbl-items-strip-search]).
+In general, there is a substantial 10% of strip searches being conducted without any reasons (see [@tbl-reasons-strip-search]).
+Commonly, strip searches are carried out by at least 1 or 2 reasons provided by the police, both at 29% (see [@tbl-reasons-strip-search]).
+As for items found during searches, at a whooping 63%, most strip search enforcement doesn't garner any suspected items (see [@tbl-items-strip-search]).
+This critically denotes how many strip searches aren't motivated by civic considerations but are moreso carried out by intuition, or personal biases that can extend to racial profiling.
+It signifies an alarming potential for misconducted strip searches that are racially or sexually motivated.
+To visualize this data, @fig-reasons-race and @fig-reasons-gender both map, respectively, the racial and gender distribution of strip searches by the total number of reasons each strip search has.
+
+Regarding @fig-reasons-race, due to the sheer number of police searches that people who are racially black and white have compared to other groups, this paper subsections the dataset on racialized police searches into two graphs.
+Both of these graphs are devoid of data from people are racially unknown or don't identify with any races--"None" or "Unknown or Legacy"--to ensure visibility.
+In specifics, graph a) of @fig-reasons-race contains only the distribution from people who racially identify as black and white, graph b) of @fig-reasons-race contains the distribution of all other minority groups--except for black people--against white people.
+This ensures data clarity between groups with higher counts of strip searches and groups with lower counts of strip searches.
+In addition, to ensure that data is proportionate to one another, instead of using `count`, the percentages for each racial group categorized by the different sums of reasons within their total number of strip searches are calculated.
+
+Based on part a) of @fig-reasons-race, the number of black people who are strip searched surpass white people on all counts of total number of reasons, except for instances where the police only has one reason to conduct strip search.
+This signifies an alarming rate of racialized strip searches against black people despite their small population make-up in Toronto (@ethnicCensus).
+For part b) of @fig-reasons-race, the percentages between other minority groups fluctuate.
+But intriguingly, compared to white people, all of other minority groups are more likely to be strip searched with only  Groups that are most commonly strip searched without any provided reasons are "Middle-Eastern", "East/Southeast Asian" and "Indigenous", at 11%, 11% and 10% each.
+
+Population-wise,
+
+```{r fig.height=4, fig.width=8}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-reasons-race
+#| fig-cap: Relationship Between Reasons for Strip Search and Race in Toronto, 2020-2021
+#| n-cols: 2
+#| fig-subcap: ["Black and White","Other Minority Groups", "new"]
+#| layout-ncol: 3
+
+#create a scatterplot to show each observation in the dataset, comparing the total reasons to racial identitiy found in each strip search
+#colour the points based on the perceived race of the captive
+#code for geom_point() referenced from https://github.com/TDonofrio62/Shooting-Occurrences-in-Toronto/blob/main/Shooting_Occurrences/outputs/paper/Shooting_Occurrences.Rmd
+
+cleaned_search_reasons_data |>
+  filter(race == "Black" | race == "White") |>
+  group_by(total_reasons, race) |>
+  count() |>
+  rename(count = n) |>
+   group_by(race) |>
+  mutate(percentage = round(count/sum(count)*100)) |>
+  ggplot(aes(x = total_reasons, y = percentage, color = race)) +
+  geom_point(aes(size = percentage)) +
+  geom_smooth() +
+  labs(x = "Total Reasons for Strip Searching",
+       y = "Percentage",
+       color = "Race"
+       ) +
+  theme_minimal() +
+  theme(legend.key.size = unit(0.5, 'cm'))
+
+cleaned_search_reasons_data |>
+  filter(race != "None" & race != "Unknown or Legacy" & race != "Black") |>
+  group_by(total_reasons, race) |>
+  count() |>
+  rename(count = n) |>
+  group_by(race) |>
+  mutate(percentage = round(count/sum(count)*100)) |>
+  ggplot(aes(x = total_reasons, y = percentage, color = race)) +
+  geom_point(aes(size = percentage)) +
+  geom_smooth() +
+  labs(x = "Total Reasons for Strip Searching",
+       y = "Percentage",
+       color = "Race"
+       ) +
+  theme_minimal() +
+  theme(legend.key.size = unit(0.5, 'cm')) 
+```
+
+```{r fig.height=3 , fig.width=8}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-reasons-gender
+#| fig-cap: Relationship Between Reasons for Strip Search and Gender in Toronto, 2020-2021
+
+cleaned_search_reasons_data |>
+  group_by(total_reasons, gender) |>
+  count() |>
+  rename(count = n) |>
+   group_by(gender) |>
+  mutate(percentage = round(count/sum(count)*100)) |>
+  ggplot(aes(x = total_reasons, y = percentage, colour = gender)) +
+  geom_point(aes(size = percentage)) +
+  geom_smooth() +
+  labs(y = "Count", x = "Total Reasons for Strip Searching", color = "Gender") +
+  theme(legend.key.size = unit(0.5, 'cm')) +
+  theme_minimal() +
+  scale_fill_viridis_d(option="cividis")
+
+```
+
+## Discussion and Conclusion
+
+'None' and unknown legacy, and unidentified
+difficult work of categorizing
+binaristic view of gender
+
+mystified, how is race sectioned and divided
+
+biased source, observation
+
+lack tornotno ethnic demofraphic data to actually compare and contrast if the perecentage of strip search is correlated to the population spread
+
+```{r fig.height=6, fig.width=8}
+#| echo: false
+#| message: false
+#| label: fig-race_strip_search_separate
+#| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
+
+cleaned_race_gender_data |>
+  group_by(race, strip_search, gender) |>
+  count() |>
+  rename(count= n) |>
+  group_by(race) |>
+  mutate(percentage = round(count/sum(count)*100)) |>
+  ggplot(mapping = aes(x = gender, fill = strip_search), data=cleaned_race_gender_data) +
+  geom_bar(position = "dodge") + 
+  labs(y = "Count", x = "Gender", fill = "Strip Search") +
+  facet_wrap(~race) + scale_fill_viridis_d(option="cividis")
+
+```
+
+Overall, the data reveals relationships between demographic factors (age, gender, and area or neighborhood) and number of MHA apprehensions reports.
+By conducting histogram graphs, We found that the number of MHA reports increases each year until 2022: in comparison, the total crime reports in Toronto decreases until 2020 (@fig-years).
+This trend of data could be a result of Covid-19 Pandemic due to the fact that many people encounter mental health issues while staying mostly at homes.
+The difference of the patterns in MHA and total crime reports also hints a potential indirect correlation with each other.
+Demographically, males in age cohort from 25 to 34 are more likely to be reported as MHA apprehensions(`r male_per`% males, `r age_per`% with age cohort from 25 to 34) as shown in @fig-groups.
+The difference between gender and age groups are not necessarily significant though are most likely to be since the large sample size.
+Closer studies such as model and power analysis should be conducted for better statistical understanding of the data sets.
+
+Additionally, an possible indirect relation between the number of MHA apprehension and corresponding area division is detected: increase in number of total crime in one division does not correlates to higher number of MHA apprehensions with the same degree within areas (@fig-merged).
+The percentages of average MHA reports out of average police reports of the years are not consistent across the divisions.
+In other words, possible bias, mediating, or moderating factors could take place in the correlation equation such as @eq-relation.
+Although the paper does not create and testify the validity of the equations and models, it provides a potential future direction for the research field.
+
+While these findings point towards a slight positive correlation between police strip search being racially inclined and gendered, particularly towards the black and female population of Toronto, it's worth noting how the data has decontextualized the reasons the police account for each strip search.
+Specifically, all reasons-- including "Injury", "Weapons", "Escape" and "Possess Evidence"--are quantified as a single unit for each arrest, without further details on the severity scale of the situation.
+In short, all of these reasons are weighed equally.
+This can very much seclude the fact that some arrests--even  are more necessary to be strip searched than others, thus skewing the nature of this paper's analysis.
+For instance, reffering back to @tbl-reasons-strip-search, the more number of reasons to account for a strip search doesn't necessarily mean that a person is bound to be strip searched.
+The number of reasons the dataset attributes to each strip search--as provided by Toronto Police Service [@torontoPolice]--can be purely projected based on the police in power's racial preconceptions or gender biases.
+This means that a person with only one reason to be strip searched may be more "necessary" to be strip searched than others.
+Nonetheless, this paper believes that the categorization of the reasons this way by the dataset somewhat confirms th
+
+That said, there are further pain-points regarding the nature of the dataset, making the analysis suggestive in nature.
+As mentioned, noting the combative nature between the Toronto police force and people from the LGBTQI+ community [@policeGender], this paper recognizes that simplifying the gender spectrum to a mere binary of males and females can disregard the struggles of non-binary people (see [@tbl-gender-strip-search]).
+This problem of 'categorization' is also problematic for the ".
+
+Additional socio-economic measures can further ascertain the validity of this dataset
+
+by summing these reasons together.
+Specifically, the severity of the reasons aren't factored within this paper analysis, which can hinder its conclusion.
+For instance, a person involved in the strip search might have caused grave injuries or possessed dangerous
+
+Although these findings point to inequity in child care access based on average income per ward, many other factors must also be taken into consideration.
+The child care data provided by the Child Service's division only contains information about licensed child care centres [@caredata].
+The Government of Ontario specifies that unlicensed home child care providers can operate with five or fewer children [@ontario].
+This may account for many of the children who are unable to obtain a space in a licensed child care facility as may family members and private sitters, but this data is unavailable through Open Data Toronto.
+The average cost of child care at each facility could help to provide more insight into equitable access but this information is also unavailable in this data set.
+
+A variable which may skew the data or produce outliers is the fact that each of Toronto's wards contain multiple neighborhoods with various social and economic considerations [@wards].
+Outliers such as Toronto Centre and Scarborough Rouge-Park for instance may be influenced by variation in income by neighborhood or higher child care supply based on commuting population and business density.
+Although average household income by ward was used in @fig-childrenspacesplot and @fig-childrenspacesscatterplot, an analysis using the median household income of each ward showed a similar trend (see Appendix: @fig-spacesmedianbywards & @fig-spacesmedianincome).
+
+## LLMs
+
+Statement on LLM usage: no Learning Management Systems (LMSs) were solicited during the writing of this paper.
+
+## Appendix {-}
+
+### Strip Search and Items Found Correlation
+
+@tbl-reasons-strip-search connotes the summary statistics how many reasons the police force attributes to each person prior to conducting strip searches and the number of items found during each search.
+This table leverages `mutate` to create three new columns on the data's percentages and counts, enabling the data to be proprotionate to one another.
+"Total Reasons" conveys the total number of reasons a strip search has, while "Items Found Percentage" conveys the percentage of items found for each total of reasons.
+"Total Searches by Total Reasons" conveys the total number of searches for each total of reasons, while "Searches Percentage" conveys the percentage of each total of reasons out of all strip searches.
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-reasons-strip-search
+#| tbl-cap: Summary Statistics of Reasons for Police Strip Searches
+
+cleaned_search_reasons_data |>
+  group_by(total_reasons, items_found) |>
+  count() |>
+  rename(count = n) |>
+  group_by(total_reasons) |>
+  mutate(percentage = round(count/sum(count)*100),
+         total_searches = sum(count),
+         percentage_total = round(total_searches/7801*100)) |>
+  kable(
+    col.names = c("Total Reasons", "Items Found", "Count", "Items Found Percentage", "Searches by Total Reasons", "Searches Percentage"),
+    booktabs = TRUE,
+    align = c("c", "c", "c", "c", "c", "c")
+  ) |> 
+  kable_styling(latex_options="scale_down")
+```
 
@@
 
-```{r}
-#| echo: false
-#| message: false
-#| label: tbl-reasons-strip-search
-#| tbl-cap: Summary Statistics of Reasons for Police Strip Searches
-
-cleaned_search_reasons_data |>
-  group_by(total_reasons, items_found) |>
-  count() |>
-  rename(count = n) |>
-  group_by(total_reasons) |>
-  mutate(percentage = round(count/sum(count)*100),
-         total_searches = sum(count),
-         percentage_total = round(total_searches/7801*100)) |>
-  kable(
-    col.names = c("Total Reasons", "Items Found", "Count", "Percentage by Reasons", "Total Searches", "Percentage by Searches"),
-    booktabs = TRUE,
-    align = c("c", "c", "c", "c", "c", "c")
-  ) |> 
-  kable_styling(latex_options="scale_down")
-```
-
-```{r fig.height=4, fig.width=8}
-#| echo: false
-#| message: false
-#| label: fig-reasons-race
-#| fig-cap: Relationship Between Reasons for Strip Search and Race in Toronto, 2020-2021
-#| fig-subcap: ["Race","Gender"]
-#| layout-ncol: 2
-
-#create a scatterplot to show each observation in the dataset, comparing the total reasons to racial identitiy found in each strip search
-#colour the points based on the perceived race of the captive
-#code for geom_point() referenced from https://github.com/TDonofrio62/Shooting-Occurrences-in-Toronto/blob/main/Shooting_Occurrences/outputs/paper/Shooting_Occurrences.Rmd
-cleaned_search_reasons_data |>
-  group_by(total_reasons, race) |>
-  count() |>
-  rename(count = n) |>
-  ggplot(aes(x = total_reasons, y = count, colour = race)) +
-  geom_line() +
-  geom_point(aes(size = count)) +
-  labs(x = "Total Reasons for Strip Searching",
-       y = "Count",
-       color = "Race"
-       ) +
-  theme_minimal() +
-  theme(legend.key.size = unit(0.5, 'cm')) +
-  scale_fill_viridis_d(option="cividis")
-
-cleaned_search_reasons_data |>
-  group_by(total_reasons, gender) |>
-  count() |>
-  rename(count = n) |>
-  ggplot(aes(x = total_reasons, y = count, colour = gender)) +
-  geom_line() +
-  geom_point(aes(size = count)) +
-  labs(y = "Count", x = "Total Reasons for Strip Searching", color = "Gender") +
-  theme(legend.key.size = unit(0.5, 'cm')) +
-  theme_minimal() +
-  scale_fill_viridis_d(option="cividis")
-
-```
-
-```{r fig.height=3, fig.width=8}
-#| echo: false
-#| message: false
-#| results: hide
-#| label: fig-reasons-gender
-#| fig-cap: Relationship Between Reasons for Strip Search and Gender in Toronto, 2020-2021
-
-cleaned_search_reasons_data |>
-  group_by(total_reasons, gender) |>
-  count() |>
-  rename(count = n) |>
-  ggplot(aes(x = total_reasons, y = count, colour = gender)) +
-  geom_line() +
-  geom_point(aes(size = count)) +
-  labs(y = "Count", x = "Total Reasons for Strip Searching", color = "Gender") +
-  theme(legend.key.size = unit(0.5, 'cm')) +
-  theme_minimal() +
-  scale_fill_viridis_d(option="cividis")
-
-```
-
-## Discussion
-
-binaristic view of gender
-
-mystified, how is race sectioned and divided
-
-biased source, observation
-
-lack tornotno ethnic demofraphic data to actually compare and contrast if the perecentage of strip search is correlated to the population spread
-
-## Conclusion
-
-## LLMs
-
-Statement on LLM usage: no Learning Management Systems (LMSs) were solicited during the writing of this paper.
-
 \newpage

### 2024-01-26 scripts/03-test_data.R
Add files via upload

@@ -1,0 +1,123 @@
+#### Preamble ####
+# Purpose: Tests cleaned data sets 
+# Author: Thomas Fox 
+# Date: 25 January 2024
+# Contact: thomas.fox@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: run 01-download_data.R and 02-data_cleaning.R
+# Any other information needed? If nothing is printed, all tests have passed
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+library(janitor)
+
+#### Load Cleaned Data Sets ####
+
+# read cleaned child care data
+test_child_care = read_csv(
+  file = "outputs/data/child_care_data.csv",
+  show_col_types = FALSE
+)
+
+# read cleaned census data
+test_census_data = read_csv(
+  file = "outputs/data/census_data.csv",
+  show_col_types = FALSE
+)
+
+# read cleaned merged child care/census data
+test_census_childcare_data = read_csv(
+  file = "outputs/data/merged_ward_data.csv",
+  show_col_types = FALSE
+)
+
+# read ward names
+test_ward_names = read_csv(
+  file = "outputs/data/ward_names.csv",
+  show_col_types = FALSE
+)
+
+
+#### Test data ####
+
+# Test cleaned child care data 
+
+if (nrow(test_child_care) != 1063) {
+  print("Size of child care data is incorrect ")
+}
+
+if (min(test_child_care$totspace) < 0) {
+  print("Negative number in child care data total space column")
+}
+
+if (min(test_child_care$ward) <= 0) {
+  print("A ward number is below the limits ward number limit (1-25)")
+}
+
+if (max(test_child_care$ward) > 25) {
+  print("A ward number is above the ward number limit (1-25)")
+}
+
+# Test cleaned child care data 
+
+if (nrow(test_census_data) != 25) {
+  print("Number of ward entries is incorrect")
+}
+
+if (min(test_census_data$avg_hh_income) < 0) {
+  print("Negative number in average house hold income column")
+}
+
+if (min(test_census_data$pop_0_to_4) < 0) {
+  print("Negative number in population 0 to 4 column")
+}
+
+if (min(test_census_data$ward) <= 0) {
+  print("A ward number is below the limits ward number limit (1-25)")
+}
+
+if (max(test_census_data$ward) > 25) {
+  print("A ward number is above the ward number limit (1-25)")
+}
+
+
+# Test merged cleaned child care and census data 
+
[…]

### 2024-01-26 scripts/03-test_data.R
Update 03-test_data.R

@@ -1,123 +1,75 @@
 #### Preamble ####
-# Purpose: Tests cleaned data sets 
-# Author: Thomas Fox 
-# Date: 25 January 2024
-# Contact: thomas.fox@mail.utoronto.ca
+# Purpose: Tests cleaned datasets 
+# Author: Quang Mai
+# Date: 24 January 2024
+# Contact: q.mai@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: run 01-download_data.R and 02-data_cleaning.R
-# Any other information needed? If nothing is printed, all tests have passed
-
+# Pre-requisites: reload 01-download_data.R and 02-data_cleaning.R
 
 #### Workspace setup ####
 library(tidyverse)
 library(dplyr)
 library(janitor)
 
-#### Load Cleaned Data Sets ####
 
-# read cleaned child care data
-test_child_care = read_csv(
-  file = "outputs/data/child_care_data.csv",
+#### Read in cleaned data ####
+
+
+# read cleaned data on race and gender and strip searches
+cleaned_race_gender_data = read_csv(
+  file = here("inputs/data/cleaned_race_gender_data.csv"),
   show_col_types = FALSE
 )
 
-# read cleaned census data
-test_census_data = read_csv(
-  file = "outputs/data/census_data.csv",
-  show_col_types = FALSE
-)
-
-# read cleaned merged child care/census data
-test_census_childcare_data = read_csv(
-  file = "outputs/data/merged_ward_data.csv",
-  show_col_types = FALSE
-)
-
-# read ward names
-test_ward_names = read_csv(
-  file = "outputs/data/ward_names.csv",
+# read cleaned data on reasons for strip search
+cleaned_search_reasons_data = read_csv(
+  file = here("inputs/data/cleaned_search_reasons_data.csv"),
   show_col_types = FALSE
 )
 
 
 #### Test data ####
 
-# Test cleaned child care data 
+test1
 
-if (nrow(test_child_care) != 1063) {
-  print("Size of child care data is incorrect ")
-}
+# Check that there are between 1 and 65276 events of police arrests and strip searches in the cleaned dataset on gender and race 
+cleaned_race_gender_data$event |> min() == 1
+cleaned_race_gender_data$event |> max() == 65276
 
-if (min(test_child_care$totspace) < 0) {
-  print("Negative number in child care data total space column")
-}
+test2
 
-if (min(test_child_care$ward) <= 0) {
-  print("A ward number is below the limits ward number limit (1-25)")
-}
[…]

### 2024-01-26 scripts/00-simulate_data.R
update and clean datasets

@@ -24,7 +24,7 @@ set.seed(56) #ensure simulated data's reproducibility
 simulated_data <-
   tibble(
     #use 1 through 65276 to represent each arrest event
-    "event" = 1:65276,
+    "id" = 1:65276,
     #randomly pick a race identified from the dataset, with replacement, 65276 times
     "race" = sample(
       x = c("White", "Unknown or Legacy", "Black", "South Asian", "Indigenous", "Middle-Eastern", "Latino", "East/Southeast Asian", "None"),
@@ -182,5 +182,4 @@ simulated_data$gender |> unique() |> length() == 3
 
 # Check that there are exactly 2 options for strip searches #
 simulated_data$strip_search |> unique() |> length() == 2
-
 

### 2024-01-26 scripts/02-data_cleaning.R

@@ -29,16 +29,16 @@ raw_police_arrests_strip_data <- read_csv("inputs/data/raw_police_arrests_strip_
 cleaned_arrests_strip_data <-
   raw_police_arrests_strip_data |>
   janitor::clean_names() |> #clean and simplify variable names 
-  select(-c(x_id, arrest_id, person_id, object_id, youth_at_arrest_under_18_years, arrest_loc_div, booked, actions_at_arrest_concealed_i, actions_at_arrest_combative, actions_at_arrest_resisted_d, actions_at_arrest_mental_inst, actions_at_arrest_assaulted_o, actions_at_arrest_cooperative)) |> # Get rid of unnecessary variables
+  select(-c(event_id, arrest_id, person_id, object_id, youth_at_arrest_under_18_years, arrest_loc_div, booked, actions_at_arrest_concealed_i, actions_at_arrest_combative, actions_at_arrest_resisted_d, actions_at_arrest_mental_inst, actions_at_arrest_assaulted_o, actions_at_arrest_cooperative)) |> # Get rid of unnecessary variables
   rename(
          reason_injury = search_reason_cause_injury, # Shorten variable names to make them readable
          reason_escape = search_reason_assist_escape,
          reason_weapons = search_reason_possess_weapons,
          reason_possess_evidence = search_reason_possess_evidence,
          race = perceived_race,
-         gender = sex
+         gender = sex,
+         id=x_id
          ) |> 
-  relocate(event_id) |> # Put the event_id column to be the first one in the dataset
   naniar::replace_with_na(replace = list(reason_injury = c("None"), # Replace the string 'None' with NA to make it easier to manipulate data later on
                                  reason_escape = c("None"),
                                  reason_weapons = c("None"),
@@ -56,13 +56,13 @@ cleaned_arrests_strip_data$strip_search[is.na(cleaned_arrests_strip_data$strip_s
 
 cleaned_race_gender_data = 
   cleaned_arrests_strip_data |>
-  select(event_id, race, gender, strip_search, items_found)
+  select(id, race, gender, strip_search, items_found)
 
 # Clean from the cleaned dataset, to keep a dataset of variables like reasons for strip searching
 # Code for as.numeric() referenced from https://stackoverflow.com/questions/31415344/using-as-numeric-with-functions-and-pipes-in-r
 cleaned_search_reasons_data = 
   cleaned_arrests_strip_data |>
[…]

### 2024-01-26 scripts/03-test_data.R

@@ -14,6 +14,11 @@ library(janitor)
 
 #### Read in cleaned data ####
 
+# read cleaned data on arrests and strip searches
+cleaned_arrests_strip_data = read_csv(
+  file = here("inputs/data/cleaned_arrests_strip_data.csv"),
+  show_col_types = FALSE
+)
 
 # read cleaned data on race and gender and strip searches
 cleaned_race_gender_data = read_csv(
@@ -30,46 +35,47 @@ cleaned_search_reasons_data = read_csv(
 
 #### Test data ####
 
-test1
 
-# Check that there are between 1 and 65276 events of police arrests and strip searches in the cleaned dataset on gender and race 
-cleaned_race_gender_data$event |> min() == 1
-cleaned_race_gender_data$event |> max() == 65276
+# Test 1: check that there are between 1 and 65276 events of police arrests and strip searches in the cleaned dataset on gender and race 
+cleaned_race_gender_data$id |> min() == 1
+cleaned_race_gender_data$id |> max() == 65276
 
-test2
 
-# Check that the dataset on race and gender of strip searches prolongs from 2020 to 2021 
-cleaned_race_gender_data$arrest_year |> min() == 2020
-cleaned_race_gender_data$arrest_year |> max() == 2021
 
-test3
+# Test 2: check that the dataset on race and gender of strip searches prolongs from 2020 to 2021 
+cleaned_arrests_strip_data $arrest_year |> min() == 2020
+cleaned_arrests_strip_data $arrest_year |> max() == 2021
 
-# Check that there are 9 races identified in the cleaned dataset on gender, race and strip search
+
+
+# Test 3: check that there are 9 races identified in the cleaned dataset on gender, race and strip search
 cleaned_race_gender_data$race |> unique() |> length() == 9
 
-test4
 
-# Check that there are 3 gender identities identified in the cleaned dataset on gender, race and strip search
+
+# Test 4: check that there are 3 gender identities identified in the cleaned dataset on gender, race and strip search
 cleaned_race_gender_data$gender |> unique() |> length() == 3
 
-test5
 
-# Check that there are exactly 2 options of "Yes" and "No" for strip searches #
+
+# Test 5: check that there are exactly 2 options of "Yes" and "No" for strip searches #
 cleaned_race_gender_data$strip_search |> unique() |> length() == 2
 
-tes6
 
-# Check that there are 7801 observations in the cleaned dataset on reasons for strip search
+
[…]

### 2024-01-26 outputs/paper/paper.qmd
final polish of graphs and paper

@@
 bibliography: references.bib
+output: 
+  pdf_document:
+    extra_dependencies: ["float"]
 ---
@@
 Foremost, this paper aims to be a form of data activism, contributing to the dialogue around police power abuse that enforces strip-search to undermine the the vulnerable populations in Toronto.
-To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using summary tables, stacked bar charts and bubble charts.
-Particularly, for each category of race and gender, the distribution bars or points would be sectioned into various categories like 'including strip searches' or 'not including' to visualize the statistics of strip searching against intersecting social identities.
+To analyze, I first map out the distribution of Toronto police arrests by race and gender and reasons for strip searching from 2020 to 2021 using R [@r] to create summary tables, double bar charts and bubble charts.
+Particularly, for each category of race, gender and reasons for strip searching, the distribution bars or points would be sectioned into various categories like 'including strip searches' or 'not including' to visualize the statistics of strip searching against intersecting social identities.
+Racially, it was found that there is indeed an overrepresented count of black people being strip searched, and minority groups are more likely to be strip searched regardless of the number of reasons for their actions at time of arrest.
+Gender-wise, females are more likely to be strip-searched with less reasons provided and despite the low severity of their actions.
+
 It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd].
@@
 
-This paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
+This paper is organized into the following sections: Data, Results and Discussion.
 First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data cleaning and extraction took place [@torontoPolice].
-Then, all the trends and findings discovered are shown in the Results section, while the Discussion section further assesses these findings.
-Ultimately, the Conclusion section summarizes the paper's main insights.
+Then, all the trends and findings discovered are shown in the Results section.
+while the Discussion section further assesses these findings and summarizes the paper's main insights.
 
@@
 
-### Overview of Data Extraction and Cleaning
+### Overview of "Police Race and Identity Based Data - Arrests and Strip Searches"
 
 Published by Toronto Police Services, this primary dataset conveys information related to all police arrests and strip searches from 2020 to 2021 (which halted refreshing on December 2, 2022).
-Originally, the dataset features 26 variables, including date of arrest, event id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, the person's action at the time of arrest, and the police's reason for strip searching along with any items found.
-Data of strip searches are lodged based on each police arrest throughout 2020 and 2021, using a unique identifier called `event_id` which identifies the nature of the strip search and the social identities of the people involved.
+Originally, the dataset features 65276 observations of police arrests and 26 variables, including date of arrest, event id, the arrested person's race, sex, and age group, arrest location, whether the arrest includes strip searching, the person's action at the time of arrest, and the police's reason for strip searching along with any items found.
+Data of strip searches are lodged based on each police arrest throughout 2020 and 2021, using a unique identifier called `id` which identifies the nature of the strip search and the social identities of the people involved.
 No other dataset was considered for the paper but this one, since it's the only dataset that communicates information on police enforcement of strip searching.
@@
 
-### Racial and Gender Identities of Police Strip Search
+### Racial and Gender Identities of Police Strip Searches
 
@@
 
-### Reasons for Police Strip Search
+### Reasons for Police Strip Searches
 
@@
 
-\newpage
-
-```{r}
+```{r fig.pos="H"}
 #| echo: false
@@
 To encapsulate, throughout 2020 and 2021, there are a total of 7801 recorded strip searches out of 65276 incidents of police arrests, which accounts for a whooping 12% of all arrests (see [@tbl-items-strip-search]).
-Excluding those whose racial identities are unknown and non-existent, by percentage, the three most affected racial groups are, in descending order, of Indigenous, black and white background, standing at 16%, 14% and 13% each (see [@tbl-race-strip-search]).
+
+\break
 
@@
 Observably, while being the the major population group in Toronto, people of white racial roots aren't the most strip searched.
-Rather, it's the racialized population of Toronto like black and Indigenous people that are being strip searched by police.
-This hints at the possibility of strip search being more racialized than it seems, targeting minority groups in Toronto.
+Rather, it's the racialized population of Toronto like black and Indigenous people that are being strip searched by police, at respectively, 16% and 14% each.
+This hints at the possibility of strip search being more racialized than it seems, targeting minority groups in Toronto due to racial biases.
+Moreover, examining @fig-race_strip_search_separate, 2434 out of 17526 black people are strip searched in Toronto from 2020 to 2021, making them overrepresented despite their small population make-up.
+Once more, this emboldens the racialized nature of police strip search.
 
-```{r fig.height = 4, fig.width = 8}
+```{r fig.height = 5, fig.width = 8, fig.pos="H"}
 #| echo: false
@@
 
+### Gendered Police Strip Searches Statistics
+
+Gender-wise, men are more likely to be strip searched compared to women when arrested, at 12% and 10% respectively (see [@tbl-gender-strip-search]).
+Though, observably, the difference is quite marginal.
+To note, people of unidentified gender identity aren't strip searched at all according to the dataset (denoted as `U` in the tables and graphs).
+
 \break
 
-### Gendered Police Strip Searches Statistics
-
-```{r}
+```{r fig.pos="H"}
 #| echo: false
@@
 ```
-
-Gender-wise, men are more likely to be strip searched compared to women, at 12% and 10% respectively (see [@tbl-gender-strip-search]).
-Though, observably, the difference is quite marginal.
-To note, people of unidentified gender identity aren't strip searched at all according to the dataset (denoted as `U` in the tables and graphs).
 
@@
   theme(legend.key.size = unit(0.5, 'cm')) +
-  labs(x = "Race",
+  labs(x = "Gender",
        y = "Percentage",
@@
 
-### Police Strip Searches and Items Found Statistics
+### Reasons for Police Strip Searches Statistics
 
@@
 
-Regarding @fig-reasons-race, due to the sheer number of police searches that people who are racially black and white have compared to other groups, this paper subsections the dataset on racialized police searches into two graphs.
-Both of these graphs are devoid of data from people are racially unknown or don't identify with any races--"None" or "Unknown or Legacy"--to ensure visibility.
-In specifics, graph a) of @fig-reasons-race contains only the distribution from people who racially identify as black and white, graph b) of @fig-reasons-race contains the distribution of all other minority groups--except for black people--against white people.
-This ensures data clarity between groups with higher counts of strip searches and groups with lower counts of strip searches.
-In addition, to ensure that data is proportionate to one another, instead of using `count`, the percentages for each racial group categorized by the different sums of reasons within their total number of strip searches are calculated.
+#### Race and Total Reasons for Strip Searching
 
-Based on part a) of @fig-reasons-race, the number of black people who are strip searched surpass white people on all counts of total number of reasons, except for instances where the police only has one reason to conduct strip search.
-This signifies an alarming rate of racialized strip searches against black people despite their small population make-up in Toronto (@ethnicCensus).
-For part b) of @fig-reasons-race, the percentages between other minority groups fluctuate.
-But intriguingly, compared to white people, all of other minority groups are more likely to be strip searched with only  Groups that are most commonly strip searched without any provided reasons are "Middle-Eastern", "East/Southeast Asian" and "Indigenous", at 11%, 11% and 10% each.
-
-Population-wise,
-
-```{r fig.height=4, fig.width=8}
+```{r fig.width=7, fig.height=3, fig.pos="H"}
 #| echo: false
@@
 #| fig-cap: Relationship Between Reasons for Strip Search and Race in Toronto, 2020-2021
-#| n-cols: 2
-#| fig-subcap: ["Black and White","Other Minority Groups", "new"]
-#| layout-ncol: 3
+#| layout-ncol: 1
+#| fig-subcap: ["Black and White","Other Minority Groups and White"]
 
@@
 
+Regarding @fig-reasons-race, due to the sheer number of police searches that people who are racially black and white have compared to other groups, this paper subsections the dataset on racialized police searches into two graphs.
+Both of these graphs are devoid of data from people are racially unknown or don't identify with any races--"None" or "Unknown or Legacy"--to ensure graph visibility.
+In specifics, graph a) of @fig-reasons-race contains only the distribution from people who racially identify as black and white, while graph b) of @fig-reasons-race contains the distribution of all other minority groups--except for black people--along with white people.
+This ensures data clarity between groups with higher counts of strip searches and groups with lower counts of strip searches.
+In addition, to ensure that data is proportionate to one another, the percentages for each racial group when subdivided into each of the different sums of reasons are calculated and used instead of `count`.
+
+Based on part a) of @fig-reasons-race, the number of black people who are strip searched surpasses white people on all counts of total number of reasons, except for instances where the police only has one reason to conduct strip search.
+This again signifies an alarming rate of racialized strip searches against black people despite their small population make-up in Toronto (@ethnicCensus).
+For part b) of @fig-reasons-race, the percentages between other minority groups fluctuate.
+But intriguingly, compared to white people, all of other minority groups are more likely to be strip searched regardless of the number of reasons at hand.
+To note, at 11%, 11% and 10% each, people of "Middle-Eastern", "East/Southeast Asian" and "Indigenous" descent are most commonly strip searched without any provided reasons.
+"South Asian" in particular surpass all other racial groups when it comes to being strip searched because of 1 or 2 reasons for their actions at the time of arrest, at 34% and 35% respectively.
+This connotes that people of "South Asian" descent are more readily strip searched by police despite having a few reasons for their actions at the time of arrest.
+Speculatively, there could be a lot of colorism at play, considering how prerelevant racial profiling is to police power abuse, where South Asian and black people generally have a darker skin color compared to all of the other races mentioned [@skinColor].
+
+#### Gender and Total Reasons for Strip Searching
+
+Similarly, for the plot that maps "Total Reasons for Strip Searching" and "Gender", to ensure that data is proportionate to one another, instead of using `count`, the percentages for each gender group when subdivided into each of the different sums of reasons are calculated.
+
+Regarding @fig-reasons-gender, intriguingly, the graph calls into question of the above data confirming how males are slightly more likely to be strip searched compared to females when they're arrested.
+However, when intersected with data like the total number of reasons attributed strip searches, there appears to be a trend of females being more frequently strip searched when there are a smaller number of reasons--0-2 reasons--regarding their actions at the time of arrests.
+Specifically, this can translate to the fact that police more readily strip-search females compared to males even if the severity of their actions may not necessitate strip searching.
+Yet again, this hints at the potential of strip searching being sexually motivated and gendered, especially when there are mostly male police officers in Toronto [@policeStats] and where strip searching is notoriously regarded as a guise for sexual assault [@stripSearchNSW].
+
 ```{r fig.height=3 , fig.width=8}
@@
   geom_smooth() +
-  labs(y = "Count", x = "Total Reasons for Strip Searching", color = "Gender") +
+  labs(y = "Percentage", x = "Total Reasons for Strip Searching", color = "Gender") +
   theme(legend.key.size = unit(0.5, 'cm')) +
@@
 
-## Discussion and Conclusion
+## Discussion
 
-'None' and unknown legacy, and unidentified
-difficult work of categorizing
-binaristic view of gender
-
-mystified, how is race sectioned and divided
-
-biased source, observation
-
-lack tornotno ethnic demofraphic data to actually compare and contrast if the perecentage of strip search is correlated to the population spread
-
-```{r fig.height=6, fig.width=8}
+```{r fig.height=7, fig.width=9, fig.pos='H'}
 #| echo: false
 #| message: false
-#| label: fig-race_strip_search_separate
+#| label: fig-race-gender-search
 #| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
 
-cleaned_race_gender_data |>
-  group_by(race, strip_search, gender) |>
-  count() |>
-  rename(count= n) |>
-  group_by(race) |>
-  mutate(percentage = round(count/sum(count)*100)) |>
-  ggplot(mapping = aes(x = gender, fill = strip_search), data=cleaned_race_gender_data) +
-  geom_bar(position = "dodge") + 
-  labs(y = "Count", x = "Gender", fill = "Strip Search") +
-  facet_wrap(~race) + scale_fill_viridis_d(option="cividis")
+  ggplot(mapping = aes(x = race, fill = strip_search), data=cleaned_race_gender_data) +
+  geom_bar() + 
+  labs(y = "Count", x = "Race", fill = "Strip Search") +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
+  facet_wrap(~gender) + scale_fill_viridis_d(option="cividis")
 
@@
 
-Overall, the data reveals relationships between demographic factors (age, gender, and area or neighborhood) and number of MHA apprehensions reports.
-By conducting histogram graphs, We found that the number of MHA reports increases each year until 2022: in comparison, the total crime reports in Toronto decreases until 2020 (@fig-years).
-This trend of data could be a result of Covid-19 Pandemic due to the fact that many people encounter mental health issues while staying mostly at homes.
-The difference of the patterns in MHA and total crime reports also hints a potential indirect correlation with each other.
-Demographically, males in age cohort from 25 to 34 are more likely to be reported as MHA apprehensions(`r male_per`% males, `r age_per`% with age cohort from 25 to 34) as shown in @fig-groups.
-The difference between gender and age groups are not necessarily significant though are most likely to be since the large sample size.
-Closer studies such as model and power analysis should be conducted for better statistical understanding of the data sets.
+### Overview of Intersectional Insights on Strip Searching
 
-Additionally, an possible indirect relation between the number of MHA apprehension and corresponding area division is detected: increase in number of total crime in one division does not correlates to higher number of MHA apprehensions with the same degree within areas (@fig-merged).
-The percentages of average MHA reports out of average police reports of the years are not consistent across the divisions.
-In other words, possible bias, mediating, or moderating factors could take place in the correlation equation such as @eq-relation.
-Although the paper does not create and testify the validity of the equations and models, it provides a potential future direction for the research field.
+Overall, the data unfolds the relationship between strip searching and social identities such as race and gender.
+Within these intersecting identities, it was found that if there is an police arrest in Toronto from 2020 to 2021, strip searching would most likely happen to an Indigenous person or a black person.
+This is a statistically improbable insight, considering the major population make-up of white people in Toronto and their huge counts of arrests as seen in past bar graphs (see [@fig-race-gender-search]).
+Out of all the minority racial groups, at 2434 people, the black population has the highest counts of strip searches.
+Regarding reasons accounted for each strip search, minority racial groups are more likely to be strip searched without any reason provided or with less reasons provided compared to those that are ethnically white.
+Gender-wise, even though arrested males prevail in terms of being strip searched (see [@fig-race-gender-search]), females are more likely to be searched with less reasons provided.
+These insights all substantiate the assumptions that Toronto police strip search is racialized and gendered, which calls for a legal and political revisions of this intrusive form of policing.
 
-While these findings point towards a slight positive correlation between police strip search being racially inclined and gendered, particularly towards the black and female population of Toronto, it's worth noting how the data has decontextualized the reasons the police account for each strip search.
-Specifically, all reasons-- including "Injury", "Weapons", "Escape" and "Possess Evidence"--are quantified as a single unit for each arrest, without further details on the severity scale of the situation.
-In short, all of these reasons are weighed equally.
-This can very much seclude the fact that some arrests--even  are more necessary to be strip searched than others, thus skewing the nature of this paper's analysis.
-For instance, reffering back to @tbl-reasons-strip-search, the more number of reasons to account for a strip search doesn't necessarily mean that a person is bound to be strip searched.
+### Loopholes and Weaknesses
+
+#### Questionable Quantification of Reasons for Strip Search
+
+While these findings point towards a positive correlation between police strip search being racially inclined and gendered, particularly towards the black and female population of Toronto, it's worth noting how the data has decontextualized the reasons the police account for each strip search.
+Specifically, each reason-- including "Injury", "Weapons", "Escape" and "Possess Evidence"--is quantified as a single, equal unit for each arrest, without other further details on the severity scale of the reason itself.
+This can very much seclude the fact that some arrests--even those having the same reason--are more necessary to be strip searched than others, thus skewing the nature of this paper's analysis.
+Moreover, referring back to @tbl-reasons-strip-search, the more number of reasons to account for a strip search doesn't necessarily mean that a person is bound to be strip searched.
 The number of reasons the dataset attributes to each strip search--as provided by Toronto Police Service [@torontoPolice]--can be purely projected based on the police in power's racial preconceptions or gender biases.
-This means that a person with only one reason to be strip searched may be more "necessary" to be strip searched than others.
-Nonetheless, this paper believes that the categorization of the reasons this way by the dataset somewhat confirms th
+This means that the reasons for each strip search has might be purely intuitive and subjective.
+
+#### Problematic Categorization of Data and Identities
 
 That said, there are further pain-points regarding the nature of the dataset, making the analysis suggestive in nature.
-As mentioned, noting the combative nature between the Toronto police force and people from the LGBTQI+ community [@policeGender], this paper recognizes that simplifying the gender spectrum to a mere binary of males and females can disregard the struggles of non-binary people (see [@tbl-gender-strip-search]).
-This problem of 'categorization' is also problematic for the ".
+As mentioned, noting the combative nature between the Toronto police force and people from the LGBTQI+ community [@policeGender], this paper recognizes that simplifying the gender spectrum to a mere binary of males and females can disregard the struggles non-binary people have against police strip searching (see [@tbl-gender-strip-search]).
+This problem of 'categorization' is also problematic for the white population in Toronto, where the dataset essentially lumps the many ethnic variations within 'whiteness' into one category.
+This thus makes the paper's analysis of the white population in relation to strip-searching one dimensional, unable to dissect how they are also acutely impacted by invasive strip searches.
+This goes the same for the "None" and "Unknown or Legacy" entries in the `race` column.
+While this is not necessarily the dataset's loophole, these entries's non-clarified data points somewhat delimit the paper analysis of how Toronto police strip searches are racialized.
 
-Additional socio-economic measures can further ascertain the validity of this dataset
+## Future Directions
 
-by summing these reasons together.
-Specifically, the severity of the reasons aren't factored within this paper analysis, which can hinder its conclusion.
-For instance, a person involved in the strip search might have caused grave injuries or possessed dangerous
+Throughout this paper, only one dataset was used to cover strip-searching in Toronto.
+Considering the subtext of COVID-19 from 2020 to 2021, where physical contact was constrained and less people can go out despite the rising police enforcement, this dataset will benefit a lot more with additional datasets on strip searching in Toronto throughout the years, especially those without COVID-19 mobility restrictions.
+This will enable a large sampling size and make the data on strip searching more well-represented.
 
-Although these findings point to inequity in child care access based on average income per ward, many other factors must also be taken into consideration.
-The child care data provided by the Child Service's division only contains information about licensed child care centres [@caredata].
-The Government of Ontario specifies that unlicensed home child care providers can operate with five or fewer children [@ontario].
-This may account for many of the children who are unable to obtain a space in a licensed child care facility as may family members and private sitters, but this data is unavailable through Open Data Toronto.
-The average cost of child care at each facility could help to provide more insight into equitable access but this information is also unavailable in this data set.
-
-A variable which may skew the data or produce outliers is the fact that each of Toronto's wards contain multiple neighborhoods with various social and economic considerations [@wards].
-Outliers such as Toronto Centre and Scarborough Rouge-Park for instance may be influenced by variation in income by neighborhood or higher child care supply based on commuting population and business density.
-Although average household income by ward was used in @fig-childrenspacesplot and @fig-childrenspacesscatterplot, an analysis using the median household income of each ward showed a similar trend (see Appendix: @fig-spacesmedianbywards & @fig-spacesmedianincome).
+Additional social variables can also further ascertain the validity of this dataset.
+Information on the people involved in each arrest like their economic background, education and health status can be some insightful variable to see how strip search affects certain social identities more than others.
+Ultimately, having these new variables, we can assess police strip search as a socio-political phenomenon, and aim to resolve its injustices and marginalization.
 
@@
 
+\newpage
+
+\appendix
+
 ## Appendix {-}
 
-### Strip Search and Items Found Correlation
+### Strip Search and Reasons for Strip Search Correlation
 
@@
 
-```{r}
+```{r fig.pos="H"}
 #| echo: false
@@
   kable_styling(latex_options="scale_down")
+
 ```
 
-```{r}
+### Strip Search and Items Found Correlation
+
+```{r fig.pos="H"}
 #| echo: false

### 2024-01-26 outputs/paper/paper.qmd
further edit graphs and final polish of paper

@@
 This hints at the possibility of strip search being more racialized than it seems, targeting minority groups in Toronto due to racial biases.
-Moreover, examining @fig-race_strip_search_separate, 2434 out of 17526 black people are strip searched in Toronto from 2020 to 2021, making them overrepresented despite their small population make-up.
+Moreover, examining tbl-race-strip-search, 2434 out of 17526 black people are strip searched in Toronto from 2020 to 2021, making them overrepresented despite their small population make-up.
 Once more, this emboldens the racialized nature of police strip search.
@@
 Regarding @fig-reasons-gender, intriguingly, the graph calls into question of the above data confirming how males are slightly more likely to be strip searched compared to females when they're arrested.
-However, when intersected with data like the total number of reasons attributed strip searches, there appears to be a trend of females being more frequently strip searched when there are a smaller number of reasons--0-2 reasons--regarding their actions at the time of arrests.
+However, when intersected with data like the total number of reasons attributed to strip searches, there appears to be a trend of females being more frequently strip searched when there are a smaller number of reasons--0-2 reasons--regarding their actions at the time of arrests.
 Specifically, this can translate to the fact that police more readily strip-search females compared to males even if the severity of their actions may not necessitate strip searching.

### 2024-01-26 outputs/paper/paper.qmd
final polish of appendix and paper

@@
 #| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
+cleaned_male_female_data <-
+  cleaned_race_gender_data |>
+  filter(gender != "U")
 
-  ggplot(mapping = aes(x = race, fill = strip_search), data=cleaned_race_gender_data) +
+  ggplot(mapping = aes(x = race, fill = strip_search), data=cleaned_male_female_data) +
   geom_bar() + 
@@
 
+@tbl-items-strip-search conveys the total number of strip searches happening in Toronto from 2020 to 2021, and whether or not items were found during each search.
+
 ```{r fig.pos="H"}

### 2024-01-26 outputs/paper/paper.qmd
polish formatting and finalize paper

@@
 title-block-banner: true
-abstract: "Police strip search has been a topic of great contention during the COVID-19 pandemic.
-The question commonly raised was how certain demographics are more likely to undergo strip searches than others, and why has this skewed power dynamics been maintained for so long.
+abstract: "Police strip search has always been a topic of great contention, where question commonly raised was how certain demographics are more likely to undergo strip searches than others.
 Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021.
@@
 This paper is organized into the following sections: Data, Results and Discussion.
-First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data cleaning and extraction took place [@torontoPolice].
+First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data extraction took place [@torontoPolice].
 Then, all the trends and findings discovered are shown in the Results section.
@@
 Thus, in order to conduct an intersectional---or gender and race---analysis of police strip searches, I have to further clean this dataset and split it into two smaller datasets.
-Additional details on data extraction and cleaning will be specified in subsequent sections.
 
@@
 Most relevant to the paper are the variables on the racial and gender identities of people who were arrested and then strip-searched.
+There are 9 unique racial groups for the `race` variables and 3 unique gender groups for the `gender` variables.
 To streamline this data, the names within the original dataset are simplified and then the dataset is split into a smaller dataset containing relevant columns ([see @tbl-cleaned_race_gender_data]).
@@
 After renaming, the big dataset was refined into containing relevant columns ([see @tbl-cleaned_search_reasons_data]).
-Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for all `reason` variables with `NA` as their values.
+Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function.
 
@@
 
-To encapsulate, throughout 2020 and 2021, there are a total of 7801 recorded strip searches out of 65276 incidents of police arrests, which accounts for a whooping 12% of all arrests (see [@tbl-items-strip-search]).
+To encapsulate, throughout 2020 and 2021, there are a total of 7801 recorded strip searches out of 65276 incidents of police arrests, which accounts for 12% of all arrests (see [@tbl-items-strip-search]).
 
@@
 Gender-wise, even though arrested males prevail in terms of being strip searched (see [@fig-race-gender-search]), females are more likely to be searched with less reasons provided.
-These insights all substantiate the assumptions that Toronto police strip search is racialized and gendered, which calls for a legal and political revisions of this intrusive form of policing.
+These insights all substantiate the assumptions that Toronto police strip searches are racialized and gendered, which calls for a legal and political revisions of this intrusive form of policing as it hinges on racial biases, profiling, social intimidation and sexual invasion.
 

### 2024-01-26 outputs/paper/paper.qmd
polish titling and finalize paper

@@
 ---
-title: "Slightly Positive Correlation Between Racialization, Gendering and Police Strip Searches in the Toronto Municipal, 2020-2021"
+title: "Positive Correlation Between Racialization, Gendering and Police Strip Searches in the Toronto Municipal, 2020-2021"
 title-block-banner: true

### 2024-01-26 outputs/paper/paper.qmd
finalize paper for submission

@@
 
-#### Questionable Quantification of Reasons for Strip Search
+#### Quantification of Reasons for Strip Search
 
@@
 This can very much seclude the fact that some arrests--even those having the same reason--are more necessary to be strip searched than others, thus skewing the nature of this paper's analysis.
-Moreover, referring back to @tbl-reasons-strip-search, the more number of reasons to account for a strip search doesn't necessarily mean that a person is bound to be strip searched.
+Moreover, referring back to @tbl-reasons-strip-search, more reasons to account for a strip search doesn't necessarily mean that a person is bound to be strip searched.
 The number of reasons the dataset attributes to each strip search--as provided by Toronto Police Service [@torontoPolice]--can be purely projected based on the police in power's racial preconceptions or gender biases.
-This means that the reasons for each strip search has might be purely intuitive and subjective.
+This means that the way the reasons are flagged out for each strip search has might be personally motivated.
 
-#### Problematic Categorization of Data and Identities
+#### Categorization of Data and Identities
 

### 2024-01-26 outputs/paper/paper.qmd
finalize paper for submission
retitle graphs, tables and finalize paper

@@
 #| label: tbl-cleaned_race_gender_data
-#| tbl-cap: Sample of Race and Gender Data against Police Arrests with Strip Searches
+#| tbl-cap: Sample of Race and Gender against Police Arrests with and without Strip Searches
 
@@
 #| label: tbl-cleaned_search_reasons_data
-#| tbl-cap: Sample of Reasons for Police Strip Searches Data
+#| tbl-cap: Sample of Reasons for Police Strip Searches
 
@@
 #| label: tbl-gender-strip-search
-#| tbl-cap: Summary Statistics of Gendered Police Strip Searches
+#| tbl-cap: Summary Statistics of Gender Identities against Police Strip Searches
 
@@
 
-```{r fig.width=7, fig.height=3, fig.pos="H"}
+```{r fig.width=7, fig.height=2.5, fig.pos="H"}
 #| echo: false
@@
 #| label: fig-race-gender-search
-#| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
+#| fig-cap: Count of Gender and Racial Identities against Police Strip Searches in Toronto
 cleaned_male_female_data <-
@@
 
-Overall, the data unfolds the relationship between strip searching and social identities such as race and gender.
+Illustrated in @fig-race-gender-search, this paper aims to interogate the relationship between strip searching and social identities such as race and gender in Toronto from 2020 to 2021.
 Within these intersecting identities, it was found that if there is an police arrest in Toronto from 2020 to 2021, strip searching would most likely happen to an Indigenous person or a black person.
@@
 The number of reasons the dataset attributes to each strip search--as provided by Toronto Police Service [@torontoPolice]--can be purely projected based on the police in power's racial preconceptions or gender biases.
-This means that the way the reasons are flagged out for each strip search has might be personally motivated.
+This means that the way the reasons are flagged out for each strip search might be personally motivated.
 

### 2024-01-26 outputs/paper/paper.qmd
resize graphs and finalize paper

@@
 
-```{r fig.pos="H"}
+```{r}
 #| echo: false
@@
 
-```{r fig.height=3, fig.width=5}
+```{r fig.height=4, fig.width=5}
 #| echo: false
@@
 
-```{r fig.width=7, fig.height=2.5, fig.pos="H"}
+```{r fig.width=7.5, fig.height=4, fig.pos="H"}
 #| echo: false
@@
 
-```{r fig.height=3 , fig.width=8}
+```{r fig.height=4.5 , fig.width=8}
 #| echo: false

### 2024-01-26 outputs/paper/paper.qmd
resize graphs and finalize paper

@@
 #referenced kable stying from https://github.com/EthanSansom/torontocitytrees/blob/main/outputs/paper/paper.Rmd
-head(cleaned_race_gender_data, 6) |>
+head(cleaned_race_gender_data, 10) |>
   kable(
@@
 
-  head(cleaned_search_reasons_data, 6) |>
+  head(cleaned_search_reasons_data, 10) |>
   select(-c(race, gender)) |>
   kable(
-    col.names = c("ID", "Reason - Injury", "Reason - Escape", "Reason - Weapons", "Reason - Has Evidence", "Items_Found", "Total Reasons"),
+    col.names = c("ID", "Reason - Injury", "Reason - Escape", "Reason - Weapons", "Reason - Has Evidence", "Items Found", "Total Reasons"),
     booktabs = TRUE,
@@
 This is due to the incongruous population spread of each racial and gender group within Toronto.
-For instance, 44.3% of Toronto population is made up of white people [@ethnicCensus], thus it's more probable that the people arrested and strip searched by the police force are ethnically "White" in the dataset (see [@tbl-race-strip-search]).
-As such, to render the data statistically proportionate to one another, `mutate()`was used to create new variables that are mostly percentage values.
+For instance, 44.3% of Toronto population is made up of white people [@ethnicCensus], thus it's more probable that the people arrested and strip searched by the police force are ethnically white in the dataset (see [@tbl-race-strip-search]).
+As such, to render the data statistically proportionate to one another, `mutate()`was used to create new columns of mostly percentage values, relating each racial and gender group to the variable being measured at hand
 

### 2024-01-26 outputs/paper/paper.qmd
fine-tune writing and finalize paper

@@
 Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021.
-Nonetheless, it's still worth looking at due the precedents of police power abuse [@covidStripSearch].
+Nonetheless, it's still worth looking at due to the precedents of police power abuse [@covidStripSearch].
 
@@
 Most relevant to the paper are the variables on the racial and gender identities of people who were arrested and then strip-searched.
-There are 9 unique racial groups for the `race` variables and 3 unique gender groups for the `gender` variables.
-To streamline this data, the names within the original dataset are simplified and then the dataset is split into a smaller dataset containing relevant columns ([see @tbl-cleaned_race_gender_data]).
-This dataset entails demographic and strip search-relevant information for each police arrest event, such as race, gender, whether the event involves a strip search, and if it is a strip search, are there any items found.
+After having their names simplified, this subset of data from the original dataset was generated, containing relevant columns on the social identities of people invovled in strip search ([see @tbl-cleaned_race_gender_data]).This dataset entails demographic and strip search-relevant information for each police arrest event, such as race, gender, whether the event involves a strip search, and any items found during serach.
+There are 9 unique racial groups for the `race` variables and 3 unique gender groups for the `gender` variables
 
-Specifically, the column 'Strip Search' signifies whether an arrest involves strip search.
+Specifically, the column 'Strip Search' signifies whether an arrest involves strip searching.
 As for the 'Items Found' column, either a `1` or a `0` will appear in a data cell if the 'Strip Search' value of an event is a `1`.
@@
 Nonetheless, the dataset only contains a binaristic view of gender, without regarding those who identify as non-binary or others, and essentially lumping all other gender identities as "Unidentified".
-This proves inconsistent against the constant public backlash agasint police for their power abuse of people belonging to the LGBTQI+ community in Toronto [@policeGender].
-This point will be further discussed below.
+This proves inconsistent against the constant public backlash agasint police for their power abuse of people belonging to the LGBTQI+ community in Toronto [@policeGender]; this point will be further discussed below.
 
@@
 Commonly, strip searches are carried out by at least 1 or 2 reasons provided by the police, both at 29% (see [@tbl-reasons-strip-search]).
-As for items found during searches, at a whooping 63%, most strip search enforcement doesn't garner any suspected items (see [@tbl-items-strip-search]).
+As for items found during searches, at a whopping 63%, most strip search enforcement doesn't garner any suspected items (see [@tbl-items-strip-search]).
 This critically denotes how many strip searches aren't motivated by civic considerations but are moreso carried out by intuition, or personal biases that can extend to racial profiling.
@@
 Both of these graphs are devoid of data from people are racially unknown or don't identify with any races--"None" or "Unknown or Legacy"--to ensure graph visibility.
-In specifics, graph a) of @fig-reasons-race contains only the distribution from people who racially identify as black and white, while graph b) of @fig-reasons-race contains the distribution of all other minority groups--except for black people--along with white people.
+In specifics, graph (a) of @fig-reasons-race contains only the distribution from people who racially identify as black and white, while graph (b) of @fig-reasons-race contains the distribution of all other minority groups--except for black people--along with white people.
 This ensures data clarity between groups with higher counts of strip searches and groups with lower counts of strip searches.
@@
 
-Based on part a) of @fig-reasons-race, the number of black people who are strip searched surpasses white people on all counts of total number of reasons, except for instances where the police only has one reason to conduct strip search.
+Based on part (a) of @fig-reasons-race, the number of black people who are strip searched surpasses white people on all counts of total number of reasons, except for instances where the police only has one reason to conduct strip search.
 This again signifies an alarming rate of racialized strip searches against black people despite their small population make-up in Toronto (@ethnicCensus).
-For part b) of @fig-reasons-race, the percentages between other minority groups fluctuate.
+For part (b) of @fig-reasons-race, the percentages between other minority groups fluctuate.
 But intriguingly, compared to white people, all of other minority groups are more likely to be strip searched regardless of the number of reasons at hand.

### 2024-01-26 outputs/paper/paper.qmd
fine-tune writing and finalize paper

@@
 
-```{r fig.height=4, fig.width=5}
+```{r fig.height=4, fig.width=5, fig.align='center'}
 #| echo: false

### 2024-01-26 outputs/paper/paper.qmd
fine-tune writing and finalize paper

@@
 title-block-banner: true
-abstract: "Police strip search has always been a topic of great contention, where question commonly raised was how certain demographics are more likely to undergo strip searches than others.
+abstract: "Police strip search has always been a topic of great contention, where the question commonly raised was how certain demographics are more likely to undergo strip searches than others.
 Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021.
-In short, this paper finds that police strip searches are slightly racialized and gendered, especially towards people of African descent and females within Toronto.
+In short, this paper finds that police strip searches are quite racialized and gendered, especially towards people of African descent and females within Toronto.
 This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities."

### 2024-01-26 outputs/paper/paper.qmd
fine-tune writing and finalize paper

@@
 
-Regarding @fig-reasons-race, due to the sheer number of police searches that people who are racially black and white have compared to other groups, this paper subsections the dataset on racialized police searches into two graphs.
+Regarding @fig-reasons-race, count-wise, since people who are racially black and white have more strip searches compared to other groups, this paper subsections the dataset on racialized police searches into two graphs.
 Both of these graphs are devoid of data from people are racially unknown or don't identify with any races--"None" or "Unknown or Legacy"--to ensure graph visibility.

### 2024-01-26 outputs/paper/paper.qmd
fine-tune writing and finalize paper

@@
 Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021.
-In short, this paper finds that police strip searches are quite racialized and gendered, especially towards people of African descent and females within Toronto.
+In short, this paper finds that police strip searches are slightly racialized and gendered, especially towards people of African descent and females within Toronto.
 This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities."
@@
 Regarding @fig-reasons-gender, intriguingly, the graph calls into question of the above data confirming how males are slightly more likely to be strip searched compared to females when they're arrested.
-However, when intersected with data like the total number of reasons attributed to strip searches, there appears to be a trend of females being more frequently strip searched when there are a smaller number of reasons--0-2 reasons--regarding their actions at the time of arrests.
+However, when intersected with data like the total number of reasons attributed to strip searches, there appears to be a different narrative.
+For strip searches that have 0 to 2 reason to be carried out, compared to the 10%, 28% and 29% of males, females prevail at 11%, 32% and 30% respectively.
+This equates to the fact that females are more frequently strip searched when there are a smaller number of reasons provided for their actions at the time of arrests.
 Specifically, this can translate to the fact that police more readily strip-search females compared to males even if the severity of their actions may not necessitate strip searching.

### 2024-01-26 outputs/paper/paper.qmd
fine-tune writing and finalize paper

@@
 Similarly, for the plot that maps "Total Reasons for Strip Searching" and "Gender", to ensure that data is proportionate to one another, instead of using `count`, the percentages for each gender group when subdivided into each of the different sums of reasons are calculated.
-
 Regarding @fig-reasons-gender, intriguingly, the graph calls into question of the above data confirming how males are slightly more likely to be strip searched compared to females when they're arrested.
@@
 This equates to the fact that females are more frequently strip searched when there are a smaller number of reasons provided for their actions at the time of arrests.
-Specifically, this can translate to the fact that police more readily strip-search females compared to males even if the severity of their actions may not necessitate strip searching.
+Specifically, police more readily strip-search females compared to males even if the severity of their actions may not necessitate strip searching.
 Yet again, this hints at the potential of strip searching being sexually motivated and gendered, especially when there are mostly male police officers in Toronto [@policeStats] and where strip searching is notoriously regarded as a guise for sexual assault [@stripSearchNSW].

### 2024-01-26 outputs/paper/paper.qmd
fine-tune writing and finalize paper

@@
 However, when intersected with data like the total number of reasons attributed to strip searches, there appears to be a different narrative.
-For strip searches that have 0 to 2 reason to be carried out, compared to the 10%, 28% and 29% of males, females prevail at 11%, 32% and 30% respectively.
+For strip searches that have 0 to 2 reasons to be carried out, compared to the 10%, 28% and 29% of males, females prevail at 11%, 32% and 30% respectively.
 This equates to the fact that females are more frequently strip searched when there are a smaller number of reasons provided for their actions at the time of arrests.

### 2024-01-26 outputs/paper/paper.qmd
fine-tune writing and finalize paper

@@
 Only one dataset was retrieved from the data package to examine the racial and gender identities of Toronto police arrests from 2020 to 2021, which is the package `Police Race and Identity Based Data - Arrests and Strip Searches` [@torontoPolice].
-Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `here` [@rHere], `ggrepel` [@rGgrepel], janitor` [@rJanitor], `kableExtra` [@rKableExtra], `naniar` [@rNaniar] and `knitr` [@rKnitr].
+Data was generated, extracted and cleaned using the open-source statistical programming language R [@r], leveraging functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `here` [@rHere], `ggrepel` [@rGgrepel], `janitor` [@rJanitor], `kableExtra` [@rKableExtra], `naniar` [@rNaniar], and `knitr` [@rKnitr].
 

### 2024-01-26 outputs/paper/paper.qmd
fine-tune writing and finalize paper

@@
 
-Specifically, the column 'Strip Search' signifies whether an arrest involves strip searching.
-As for the 'Items Found' column, either a `1` or a `0` will appear in a data cell if the 'Strip Search' value of an event is a `1`.
+Specifically, the column "Strip Search" signifies whether an arrest involves strip searching.
+As for the "Items Found" column, either a `1` or a `0` will appear in a data cell if the "Strip Search" value of an event is a `1`.
 This is because the items found within a strip search is only available when a police arrest does involve strip searching.
